### 2012-05-09 Thesis - Presentation/draft1.tex
Wednesday MAY 8th 2012 13:53

\documentclass[a4paper,10pt]{report}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage{abssymb}
%opening
\title{Thesis Draft 1}
\author{Anupam Prasad Vedurmudi}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Scattering Theory}

\subsection{Formal Scattering Theory}
The spectrum of any

We define the Hamiltonian operator as,
\begin{equation}
 H = H_0+V\\
\end{equation}
\noindent Where, $H_0$ is the free Hamiltonian given by,

\begin{equation}
 H_0=-\frac{\Delta}{2}
\end{equation}
\noindent $V$ is the potential.

\noindent The Green's function is formally defined as,
\begin{eqnarray}\label{GFdefinition}
 G\left(E+i\epsilon\right) &=\left(E+i\epsilon-H\right)^{-1}\\
 G_0\left(E+i\epsilon\right) &=\left(E+i\epsilon-H_0\right)^{-1}
\end{eqnarray}

\noindent The Lippmann-Schwinger equation for the Green's function is given by,
\begin{equation}\label{LS1}
 G=G_0+G_0VG
\end{equation}

\noindent Subsequently, the Born series for $G$ is obtained by recursively substituting
the $(n-1)^{th}$ order approximation for $G$ in the R.H.S of \eqref{LS1} to get the $n^{th}$
order term. The $0^{th}$ order approximation is the free Green's function $G_0$, the $1^{st}$
order approximation (the Born approximation) is given by,

\begin{equation}\label{BornApproximation}
 G=G_0+G_0VG_0
\end{equation}

The Born series can thus be written as,

\begin{equation}\label{BornSeriesExpanded}
 G=G_0+G_0VG_0+G_0VG_0VG_0+\ldots
\end{equation}

\noindent Or more compactly,
\begin{equation}\label{BornSeriesCompact}
 G=\displaystyle\sum_{n=0}^{\infty}G_0\left(VG_0\right)^n
\end{equation}

\section{Finite Element Method}
The Finite Element Method (FEM) is a numerical technique used to solve differential equations.
The method exploits the locality of the differential operator. The solution space is divided into
subspaces. These subspaces take the form of a mesh, i.e., a collection of simplices - triangles
in 2D, tetrahedra in 3D, etc. The function is then approximated using smooth basis functions with support in these subspaces.
These basis functions with finite support are referred to as finite elements.

Let $\psi$ be a solution to a differential equation. We denote the basis functions by $\phi_n$,
\begin{equation}\label{femfirst}
 \psi(\mathbf{x})\simeq\widetilde{\psi}(\mathbf{x})=\displaystyle\sum_n^N c_n\phi_n(\mathbf{x})
\end{equation}

A feature of FEM is that the basis functions have a finite overlap i.e., they are not orthogonal.
The overlap matrix in general has a block-diagonal form. The overlap matrix is denoted by,
\begin{equation}\label{overlap}
 S_{nm}=\int d^3x\phi_n(\mathbf{x})\phi_m(\mathbf{x})
\end{equation}

We also define the matrix,
\begin{equation}\label{diff_overlap}
 B_{nm}=\int d^3x\nabla\phi_n(\mathbf{x}).\nabla\phi_m(\mathbf{x})
\end{equation}

\subsection{Boundary Conditions}
In order to solve the differential equations we must also specify appropriate boundary
conditions. Let $\Omega$ be the solution space. We denote its boundary by $\partial\Omega$.

\subsubsection{Dirichlet Boundary Condition}
The Dirichlet boundary condition requires that the function has a known value on the boundary.
\begin{equation}\label{dirichlet1}
 \psi(\mathbf{x})=f(\mathbf{x}), \mbox{ }\forall\mathbf{x}\in \partial\Omega
\end{equation}
Where, $f$ is a known function on $\Omega$.

\subsubsection{Neumann Boundary Condition}
The Neumann boundary condition requires that the gradient of the function
normal to the boundary is fixed to a known function. That is,
\begin{equation}\label{neumann1}
 \frac{\partial\psi(\mathbf{x})}{\partial\mathbf{n}}\equiv\nabla\psi(\mathbf{x}).\mathbf{n} = f(\mathbf{x}), \mbox{ }\forall\mathbf{x}\in \partial\Omega
\end{equation}

\subsubsection{Mixed Boundary Condition}
It is also possible to have a boundary such that the function has a Dirichlet boundary
condition on one part (denoted by $\partial\Omega_D$) and a Neumann boundary condition
on the other ($\partial\Omega_N$).

\subsection{Python Implementation}
\section{FEM Formulation of Quantum Mechanics}
The Schrödinger equation for a wave function $\psi$ is given by,

\begin{equation}\label{schroedinger1}
 -\frac{\Delta}{2}\psi+V(\mathbf{x})\psi=E\psi
\end{equation}

\noindent Using the FEM approximation defined in \eqref{femfirst},
\begin{equation}\label{schroedinger}
 -\frac{\Delta}{2}\displaystyle\sum_n^N c_n\phi_n(\mathbf{x})+\displaystyle\sum_n^N c_n V(\mathbf{x})\phi_n(\mathbf{x})=E\displaystyle\sum_n^N c_n\phi_n(\mathbf{x})
\end{equation}
\noindent We get the weak formulation of the problem by multiplying both sides by $\phi_m(\mathbf{x})$ and integrating over. After integrating by parts, the kinetic term gives us
the following matrix element,
\begin{equation}\label{kineticterm}
 \int d^3x\phi_m(\mathbf{x})\Delta\phi_n(\mathbf{x}) = -\int d^3x\nabla\phi_m(\mathbf{x}).\nabla\phi_n(\mathbf{x}) = -B_{mn}
\end{equation}

\noindent The potential term is calculated from the known functions $\phi_n$ and is defined as,
\begin{equation}\label{potentialterm}
 \int d^3xV(\mathbf{x})\phi_m(\mathbf{x})\phi_n(\mathbf{x}) = V_{mn}
\end{equation}

\noindent Finally, the R.H.S is given by,
\begin{equation}
 E\int d^3x\phi_m(\mathbf{x})\phi_n(\mathbf{x})=ES_{mn}
\end{equation}

\noindent We define the coefficient vector $\mathbf{\underline{c}}=[c_1,c_2,\ldots,c_N]^T$. The Schrödinger equation reduces to 
the following eigenvalue problem,
\begin{equation}\label{schroedingerfem}
  H\mathbf{\underline{c}}=ES\mathbf{\underline{c}}
\end{equation}
\noindent Where, $H=\left(\frac{B}{2}+V\right)$

The inner product of two wavefunctions $\phi$ and $\chi$ is given by,
\begin{eqnarray}\label{innerproduct}
 \int d^3x \chi^*(\mathbf{x})\psi(\mathbf{x}) &=& \displaystyle\sum_{m,n}^{N}c_md_n\int d^3x \phi_m(\mathbf{x})\phi_n(\mathbf{x})\\
 &=& \mathbf{c}^{\dagger}S\mathbf{d}
\end{eqnarray}
\noindent Where, $\mathbf{c}^{\dagger}=(\mathbf{c}^*)^T$.

\subsection{Momentum Eigenstates}
We now want to find an FEM representation for free eigenstate of momentum $\mathbf{k}$ (energy,
$E_k=|\mathbf{k}|^2/2$). 
It is most convenient to do so by defining a ``cosine eigenvector'',
\begin{equation}\label{cosevec}
 |\mathbf{k}\rangle = exp(-i\mathbf{k}.\mathbf{x}) + exp(i\mathbf{k}.\mathbf{x})
\end{equation}
These can be obtained by ``solving'' the free Schrödinger equation on a grid with open
boundaries, i.e., by solving the eigenvalue problem,

\begin{equation}\label{freeschroedinger}
 H_0\psi=\lambda S\psi
\end{equation}
for $\psi$.

$|\mathbf{k}\rangle$ has the following properties,
\begin{eqnarray}
 \langle\mathbf{k}|S|\mathbf{k}\rangle &=&\int_{\Omega} d^3x\\
 \langle\mathbf{k}^{\prime}|S|\mathbf{k}\rangle &\simeq& 0
\end{eqnarray}
Here, $\int_{\Omega} d^3x$ is the volume of the solution space.

\subsection{Born Series - FEM Formulation}
We calculate the FEM representation of the Greens Function using \eqref{schroedingerfem}. We require that,
\begin{equation}\label{greenfemreq}
 G_{fem}(z)\psi=(z-E_0)^{-1}S\psi
\end{equation}
\noindent Where, $\psi$ is an eigenstate of $H$ of energy $E_0$. Thus, we define the Green's function as,
\begin{equation}\label{greenfemdef}
 G_{fem}(z)=\left(z\mathbf{I}-HS^{-1}\right)^{-1}S
\end{equation}
\noindent Where, $\mathbf{I}$ is the identity matrix. The free Greens function is defined similarly by setting
$H=B/2$. We will ultimately be testing the convergence of, $\langle\mathbf{k}|G_{fem}(z)|\mathbf{k}\rangle$.
For later convenience, the potential matrix is also suitably modified as,
\begin{equation}\label{potmat}
 V_{fem}=S^{-1}VS^{-1}
\end{equation}

\subsection{Modification of the Hamiltonian}
In this section we use the techniques outlined in \cite{KukPom76}. It this paper, the Born
series is rearranged by using orthogonally projecting pseudopotentials (OPP). It is then
proved that the rearranged Born series converges for all negative and small positive energies
even if the system contains bound states.

Let the bound states (indexed by $a$) of the Hamiltonian $H$ be denoted by $|\psi_a\rangle$.
The potential is modified by the operator $\Gamma$ given by,
\begin{equation}\label{potmod}
  \Gamma=\lambda\displaystyle\sum_a |\psi_a\rangle\langle\psi_a|
\end{equation}
\noindent The rate of convergence of the Born series is controlled by the parameter $\lambda$.

The modified Hamiltonian and the modified free Hamiltonian are denoted by,

\begin{eqnarray}
 \widetilde{H}&=&H+\Gamma\\
 \widetilde{H}_0&=&H_0+\Gamma
\end{eqnarray}

The modified Greens functions $\widetilde{G}$ and $\widetilde{G}_0$ are defined similarly as
in \eqref{GFdefinition}. The modified Born Series for $\widetilde{G}$ is given by,

\begin{equation}\label{BornSeriesModCompact}
 \widetilde{G}=\displaystyle\sum_{n=0}^{\infty}\widetilde{G}_0\left(V\widetilde{G}_0\right)^n
\end{equation}

\noindent Physically, this means that the modified Hamiltonian has the same continuous spectrum but the discrete
spectrum has been shifted by $\lambda$.

\section{Toy Problem - Scattering in 1D}
We now use the above methods in an example with scattering off a potential in 1D. Due to the simplicity of the problem,
it is possible to explicitly calculate the exact Green's function by inverting the appropriate matrix according the formal
definition given in \eqref{GFdefinition}. We will examine the problem in two cases: 
\begin{itemize}
 \item In the first case, we have convergence with both the modified and unmodified Hamiltonian but at a faster rate with the modification.
 \item In the second case, the modified hamiltonian gives a convergent Born series whereas the unmodified one diverges. 
\end{itemize}

Our box is defined as an FEM axis from $x=0$ to $x=10$ with $n=400$ discretization points, 
of order $40$ and with open boundaries. This means that the axis is divided into 400 
parts and each part has functions (elements) upto order $40$. Order $1$ being a straight
line. As an example, we show an axis, $0\leq x\leq 4$ with $n=4$ discretization points, open
boundaries and order $1$.

\begin{figure}[hm]
\centering
\includegraphics[width=225pt, height=150pt]{femexample1.png}
\caption[\textwidth]{FEM axis. n=4, order=1}
\end{figure}
The different elements are distinguished by colour.

The ``cosine eigenvectors'' are computed using \eqref{cosevec} and \eqref{freeschroedinger}.
The energies of the free wavefunctions are $E_n = n^2\pi^2/200$ for $n=0,\ldots,19$. The factor
of $200$ is a result of the discretization. The Green's function, $G(z)$ is calculated for $z=E_n+i\epsilon$,
where, $\epsilon$ is a small positive number defined based on the problem. The purpose of epsilon is to avoid 
the singularity on the positive real axis.

\subsection{First Case}
In the first case, we define the potential as a finite well of strength $0.1$ with support in $4 \leq x \leq 6$,
\begin{equation}\label{potentialdef1}
V = \begin{cases}
-0.1 &\mbox{} 4\leq x\leq 6\\
0 &\mbox{ otherwise}
\end{cases} 
\end{equation}

\begin{figure}[ht]
\centering
\includegraphics[width=225pt, height=150pt]{potential1d1.png}
\caption[\textwidth]{Test Potential 1}
% 
% \centering
% \includegraphics[width=225pt, height=150pt]{potential1d2.png}
% \caption[\textwidth]{Test Potential 2}
\end{figure}

The Born Series is calculated for $15$ iterations. The Hamiltonian is modified using the operator defined in \eqref{potmod} with
$\lambda=1$.  We also set $\epsilon=0.1$.

We see that the Born Series converges to the exact value in both cases but that the modified case converges in fewer iterations.

\subsection{Second Case}
In the second case, we define the potential as, 
\begin{equation}\label{potentialdef2}
V = -200\mathtt{exp}(-(x-3.5)^2/0.5)-200\mathtt{exp}(-(x-6.5)^2/0.5)
\end{equation}

\begin{figure}[hb]
\centering
\includegraphics[width=225pt, height=150pt]{potential1d2.png}
\caption[\textwidth]{Test Potential 2}
\end{figure}

The born series is calculated for $15$ iterations. We choose $\lambda=700$ and $\epsilon=10^{-8}$.

We see that the Born Series in the modified case converges to the exact value but diverges in the original case.

\bibliographystyle{plain}
\bibliography{draft1.bib}
\end{document}

### 2012-05-09 python/bornseriestest5_draft2.py

@@ -30,12 +30,12 @@ V2=np.zeros([y.len(),y.len()])
 
 Gam=np.zeros([y.len(),y.len()])
 
-V0=10.
+V0=0.1
 
 iter1=0
 
 t=np.linspace(0,10,1001)
-tpot=potential(t,'gaussian',np.array([3.,4.,V0]))+potential(t,'gaussian',np.array([6.,7.,V0]))
+tpot=potential(t,'fwell',np.array([4.,6.,V0]))
 
 plt.plot(t,tpot)
 plt.show()
@@ -43,14 +43,14 @@ plt.show()
 
 for e in y.e:
         b=e.matrix('d|d')
-        v=e.matrix('gaussian', np.array([6.,7.,V0]))
-        v2=e.matrix('gaussian', np.array([3.,4,V0]))
+        v=e.matrix('fwell', np.array([4.,6.,V0]))
+#        v2=e.matrix('gaussian', np.array([3.,4,V0]))
         iter2=int(np.sqrt(np.size(b)))
         for k1 in range(0,iter2):
                 for k2 in range(0,iter2):
                         B[iter1+k1,iter1+k2]=B[iter1+k1,iter1+k2]+b[k1][k2]
                         V1[iter1+k1,iter1+k2]=V1[iter1+k1,iter1+k2]+v[k1][k2]
-                        V2[iter1+k1,iter1+k2]=V2[iter1+k1,iter1+k2]+v2[k1][k2]
+#                        V2[iter1+k1,iter1+k2]=V2[iter1+k1,iter1+k2]+v2[k1][k2]
         iter1=iter1+iter2-1
 
 [evals,evecs]=la.eig(B/2. + V1 + V2,y.overlap())
@@ -68,7 +68,7 @@ evals=evals[perm]
 evecs=evecs[:,perm]
 
 #Normalization and Potential Modification
-Lambda=700.
+Lambda=1.
 for k in range(0,y.len()):
 	cosnorm=np.sqrt(2 * y.FEM_InnerProduct(cos_evecs[:,k],cos_evecs[:,k]) / (ub-lb))
 	cos_evecs[:,k]=cos_evecs[:,k] / cosnorm
@@ -80,7 +80,7 @@ for k in range(0,y.len()):
         Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
 
 niter=15
-eps=1e-8j
+eps=1e-1j
 
 #Free Green's Operator
 n0=0
[…]

### 2012-05-09 python/test5plot/plotresults.py

@@ -2,8 +2,8 @@ import pickle as pickle
 import numpy as np
 import matplotlib.pyplot as plt
 
-f1=open('test5results','r')
-f2=open('test5results_unmodified','r')
+f1=open('test5results3','r')
+f2=open('test5results3_unmodified','r')
 
 y1=pickle.load(f1)
 y2=pickle.load(f2)
@@ -14,7 +14,7 @@ plotval2=abs(y2[3:,:])
 n=np.linspace(0,14,15);
 
 for k in range(4):
-	mom_index=k * 6
+	mom_index=k+4
 	axlim1=min(plotval1[:,mom_index])
 	axlim2=1.01*max(plotval1[:, mom_index])
 	yexact=abs(y1[0,mom_index]) * np.ones(15)
@@ -25,13 +25,17 @@ for k in range(4):
 plt.show()
 
 
+#print plotval2[:,0]
+#exit('here')
 for k in range(4):
-	mom_index=k * 1
-	axlim1=min(np.log10(plotval2[:4,mom_index]))
-	axlim2=max(np.log10(plotval2[:4,mom_index]))
-	yexact=abs(y2[0,mom_index]) * np.ones(4)
+	mom_index=k+4
+	axlim1=min(plotval2[:,mom_index])
+	axlim2=max(plotval2[:,mom_index])
+	yexact=abs(y2[0,mom_index]) * np.ones(15)
 	plt.subplot(2,2,k+1)
-	plt.plot(n[:4],np.log10(plotval2[:4,mom_index]),'ro',n[:4],np.log10(yexact),'b--')
+#	plt.plot(n[:4],np.log10(plotval2[:4,mom_index]),'ro',n[:4],np.log10(yexact),'b--')
+#	plt.axis([-1, 15, axlim1, axlim2])
+	plt.plot(n,plotval2[:,mom_index],'ro',n,yexact,'b--')
 	plt.axis([-1, 15, axlim1, axlim2])
 
 plt.show()

### 2012-05-09 Thesis - Presentation/draft1.tex
Wed MAY 9th 2012 23:08

@@
 %opening
-\title{Thesis Draft 1}
+\title{Draft 1}
 \author{Anupam Prasad Vedurmudi}
@@
 \subsection{Python Implementation}
-\section{FEM Formulation of Quantum Mechanics}
+\section{FEM in Quantum Mechanics}
 The Schrödinger equation for a wave function $\psi$ is given by,
@@
 it is possible to explicitly calculate the exact Green's function by inverting the appropriate matrix according the formal
-definition given in \eqref{GFdefinition}. We will examine the problem in two cases: 
+definition given in \eqref{GFdefinition}. We will examine the problem in three cases: 
 \begin{itemize}
  \item In the first case, we have convergence with both the modified and unmodified Hamiltonian but at a faster rate with the modification.
- \item In the second case, the modified hamiltonian gives a convergent Born series whereas the unmodified one diverges. 
+ \item In the second case, the unmodified case converges for certain energies and diverges for others.With the modification, we always have
+       faster convergence.
+ \item In the third case, the modified hamiltonian gives a convergent Born series whereas the unmodified one diverges. 
 \end{itemize}
@@
 
-We see that the Born Series converges to the exact value in both cases but that the modified case converges in fewer iterations.
+We now plot the elements of the Born Series against the number of iterations. In order
+to compare the modified and unmodified cases, we plot them side by side with the modified case on the left and the unmodified one
+on the right. We follow this convention in the other two cases as well.
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series1a.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series1b.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series1c.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+We see that the Born Series converges to the exact value in both cases. It is also to be noted that the 
+.
 
@@
 \begin{equation}\label{potentialdef2}
-V = -200\mathtt{exp}(-(x-3.5)^2/0.5)-200\mathtt{exp}(-(x-6.5)^2/0.5)
+V = \begin{cases}
+-0.5 &\mbox{} 4\leq x\leq 6\\
+0 &\mbox{ otherwise}
+\end{cases} 
 \end{equation}
@@
 
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series2a.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series2b.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series2c.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+We see that for the original Hamiltonian, the Born series converges in some cases but doesn't in others. However,
+with the modification, we always have convergence at a faster rate.
+\subsection{Third Case}
+\begin{equation}\label{potentialdef2}
+V = -200\mathtt{exp}(-(x-3.5)^2/0.5)-200\mathtt{exp}(-(x-6.5)^2/0.5)
+\end{equation}
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series3a.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series3b.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}[lht]
+\includegraphics[width=360pt, height=200pt]{series3c.png}
+%\caption[\textwidth]{First Case}
+\end{figure}
+\end{center}
+
 We see that the Born Series in the modified case converges to the exact value but diverges in the original case.
-
 \bibliographystyle{plain}

### 2012-05-09 python/bornseriestest5_draft2.py

@@ -30,7 +30,7 @@ V2=np.zeros([y.len(),y.len()])
 
 Gam=np.zeros([y.len(),y.len()])
 
-V0=0.1
+V0=.1
 
 iter1=0
 
@@ -137,6 +137,6 @@ for k in range(nenergy):
 
 #print Lambda, niter
 
-f=open('test5plot/test5results3','w')
+f=open('test5plot/test5results2','w')
 pickle.dump(store1,f)
 

### 2012-05-09 python/test5plot/plotresults.py

@@ -8,36 +8,58 @@ f2=open('test5results3_unmodified','r')
 y1=pickle.load(f1)
 y2=pickle.load(f2)
 
+
 plotval1=abs(y1[3:,:])
 plotval2=abs(y2[3:,:])
 
 n=np.linspace(0,14,15);
 
-for k in range(4):
-	mom_index=k+4
-	axlim1=min(plotval1[:,mom_index])
+#for k in range(4):
+#	mom_index=k+4
+#	axlim1=min(plotval1[:,mom_index])
+#	axlim2=1.01*max(plotval1[:, mom_index])
+#	yexact=abs(y1[0,mom_index]) * np.ones(15)
+#	plt.subplot(2,2,k+1)
+#	plt.plot(n,plotval1[:,mom_index],'ro',n,yexact,'b--')
+#	plt.axis([-1, 15, axlim1, axlim2])
+#
+#plt.show()
+#
+#
+##print plotval2[:,0]
+##exit('here')
+#for k in range(4):
+#	mom_index=k+4
+#	axlim1=min(plotval2[:,mom_index])
+#	axlim2=max(plotval2[:,mom_index])
+#	yexact=abs(y2[0,mom_index]) * np.ones(15)
+#	plt.subplot(2,2,k+1)
+#	plt.plot(n,plotval2[:,mom_index],'ro',n,yexact,'b--')
+#	plt.axis([-1, 15, axlim1, axlim2])
+#
+#plt.show()
+
+for k in range(1):
+	mom_index=k+8
+
+	axlim1=.9*min(plotval1[:,mom_index])
 	axlim2=1.01*max(plotval1[:, mom_index])
 	yexact=abs(y1[0,mom_index]) * np.ones(15)
-	plt.subplot(2,2,k+1)
+	plt.subplot(1,2,k+1)
 	plt.plot(n,plotval1[:,mom_index],'ro',n,yexact,'b--')
-	plt.axis([-1, 15, axlim1, axlim2])
+	plt.plot(n,plotval1[:,mom_index],'r--')
+#	plt.axis([-1, 15, axlim1, axlim2])
 
-plt.show()
-
-
-#print plotval2[:,0]
-#exit('here')
-for k in range(4):
-	mom_index=k+4
 	axlim1=min(plotval2[:,mom_index])
 	axlim2=max(plotval2[:,mom_index])
 	yexact=abs(y2[0,mom_index]) * np.ones(15)
[…]

### 2012-05-11 Thesis - Presentation/draft1.tex
FRI May 11th 2012 12:21

@@
 \end{equation}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=225pt, height=150pt]{potential1d1.png}
-\caption[\textwidth]{Test Potential 1}
-% 
-% \centering
-% \includegraphics[width=225pt, height=150pt]{potential1d2.png}
-% \caption[\textwidth]{Test Potential 2}
+\begin{center}
+\begin{figure}[htb]
+\includegraphics[width=360pt, height=175pt]{series1a.png}
+\includegraphics[width=360pt, height=175pt]{series1b.png}
+\includegraphics[width=360pt, height=175pt]{series1c.png}
+\caption[\textwidth]{First Case - Weak Finite Well}
 \end{figure}
-
+\end{center}
 The Born Series is calculated for $15$ iterations. The Hamiltonian is modified using the operator defined in \eqref{potmod} with
@@
 on the right. We follow this convention in the other two cases as well.
-
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series1a.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
-
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series1b.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
-
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series1c.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
-
-We see that the Born Series converges to the exact value in both cases. It is also to be noted that the 
-.
 
@@
 \end{equation}
-
-\begin{figure}[hb]
-\centering
-\includegraphics[width=225pt, height=150pt]{potential1d2.png}
-\caption[\textwidth]{Test Potential 2}
-\end{figure}
+i.e., a finite well of strength $0.5$ with support in $4 \leq x \leq 6$,
 
@@
 \subsection{Third Case}
-\begin{equation}\label{potentialdef2}
+\begin{equation}\label{potentialdef3}
 V = -200\mathtt{exp}(-(x-3.5)^2/0.5)-200\mathtt{exp}(-(x-6.5)^2/0.5)
 \end{equation}
+
+\begin{figure}[hb]
+\centering
+\includegraphics[width=225pt, height=150pt]{potential1d2.png}
+\caption[\textwidth]{Test Potential 3}
+\end{figure}
 

### 2012-05-11 python/bornseriestest5_draft3.py

@@ -30,7 +30,7 @@ V2=np.zeros([y.len(),y.len()])
 
 Gam=np.zeros([y.len(),y.len()])
 
-V0=10
+V0=.1
 
 iter1=0
 
@@ -68,7 +68,7 @@ evals=evals[perm]
 evecs=evecs[:,perm]
 
 #Normalization and Potential Modification
-Lambda=100.
+Lambda=1.
 for k in range(0,y.len()):
 	cosnorm=np.sqrt(2 * y.FEM_InnerProduct(cos_evecs[:,k],cos_evecs[:,k]) / (ub-lb))
 	cos_evecs[:,k]=cos_evecs[:,k] / cosnorm
@@ -77,8 +77,7 @@ for k in range(0,y.len()):
 	evecs[:,k]=evecs[:,k] / evecsnorm
 
 	#Potential Modification
-	if evals[k]<0:
-	        Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
+        Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
 
 niter=15
 eps=1e-1j
@@ -138,6 +137,6 @@ for k in range(nenergy):
 
 #print Lambda, niter
 
-#f=open('test5plot/test5results2','w')
-#pickle.dump(store1,f)
+f=open('test5plot/test5results2','w')
+pickle.dump(store1,f)
 

### 2012-05-11 python/bornserieswtf.py

@@ -30,7 +30,7 @@ V2=np.zeros([y.len(),y.len()])
 
 Gam=np.zeros([y.len(),y.len()])
 
-V0=10
+V0=20
 
 iter1=0
 
@@ -68,7 +68,7 @@ evals=evals[perm]
 evecs=evecs[:,perm]
 
 #Normalization and Potential Modification
-Lambda=100.
+#Lambda=100.
 for k in range(0,y.len()):
 	cosnorm=np.sqrt(2 * y.FEM_InnerProduct(cos_evecs[:,k],cos_evecs[:,k]) / (ub-lb))
 	cos_evecs[:,k]=cos_evecs[:,k] / cosnorm
@@ -77,8 +77,8 @@ for k in range(0,y.len()):
 	evecs[:,k]=evecs[:,k] / evecsnorm
 
 	#Potential Modification
-	if evals[k]<0:
-	        Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
+#	if evals[k]<0:
+#	        Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
 
 niter=15
 eps=1e-1j
@@ -86,32 +86,30 @@ eps=1e-1j
 #Free Green's Operator
 n0=0
 nenergy=20
-store1=np.zeros([niter+3,nenergy])+0j
+store1=np.zeros([niter+2,nenergy])+0j
 
-Bmod=np.dot(B/2. + Lambda*Gam,y.overlap_inv())
-Bmod1=np.dot(B/2. + V1 + V2 + Lambda*Gam,y.overlap_inv())
+Bmod=np.dot(B/2.,y.overlap_inv())
 B_orig=np.dot(B/2. + V1 + V2,y.overlap_inv())
 
 Vtemp=np.dot(V1+V2,y.overlap_inv())
 Vmod=np.dot(y.overlap_inv(),Vtemp)
 
-Gamtemp=np.dot(Lambda*Gam,y.overlap_inv())
-Gam_ver=np.dot(y.overlap_inv(),Gamtemp)
 Tmat=Vmod
 
 
 for k in range(nenergy):
+	for l in range(0,y.len()):
+		if evals[l]<0:	Gam=Gam+y.FEM_Outer(evecs[:,l],evecs[:,l])/(cos_evals[k+n0]-evals[l])
[…]

### 2012-05-13 python/bornseriestest5_draft3.py
Sunday MAY 13 2012 20:43

@@ -30,7 +30,7 @@ V2=np.zeros([y.len(),y.len()])
 
 Gam=np.zeros([y.len(),y.len()])
 
-V0=.1
+V0=1
 
 iter1=0
 
@@ -68,7 +68,7 @@ evals=evals[perm]
 evecs=evecs[:,perm]
 
 #Normalization and Potential Modification
-Lambda=1.
+Lambda=1000.
 for k in range(0,y.len()):
 	cosnorm=np.sqrt(2 * y.FEM_InnerProduct(cos_evecs[:,k],cos_evecs[:,k]) / (ub-lb))
 	cos_evecs[:,k]=cos_evecs[:,k] / cosnorm
@@ -77,10 +77,11 @@ for k in range(0,y.len()):
 	evecs[:,k]=evecs[:,k] / evecsnorm
 
 	#Potential Modification
-        Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
+	if evals[k] < 0:
+	        Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
 
-niter=15
-eps=1e-1j
+niter=20
+eps=1e-5j
 
 #Free Green's Operator
 n0=0
@@ -97,7 +98,6 @@ Vmod=np.dot(y.overlap_inv(),Vtemp)
 Gamtemp=np.dot(Lambda*Gam,y.overlap_inv())
 Gam_ver=np.dot(y.overlap_inv(),Gamtemp)
 Tmat=Vmod
-
 
 for k in range(nenergy):
 	Emat=np.eye(y.len()) * (cos_evals[k+n0]+eps)
@@ -124,9 +124,19 @@ for k in range(nenergy):
 		vec1=np.dot(Gmat,cos_evecs[:,k+n0])
 		store1[l+3,k]=np.dot(cos_evecs[:,k+n0],vec1)
 
-	Gver_temp=la.inv(np.eye(y.len())+np.dot(Gmat,Gam_ver))
+	reverse_transform=np.zeros([y.len(),y.len()])
+
+	for l in range(y.len()):
+		if evals[l]<0:
+			reverse_transform=reverse_transform+y.FEM_Outer(evecs[:,l],evecs[:,l])/(cos_evals[k+n0]+eps-evals[l])
+		else: break
+#	Gver_temp=la.inv(np.eye(y.len())+np.dot(Gmat,Gam_ver))
+#	Gver=np.dot(Gver_temp,Gmat)
+
[…]

### 2012-05-13 python/bornserieswtf.py

@@ -28,18 +28,9 @@ B=np.zeros([y.len(),y.len()])
 V1=np.zeros([y.len(),y.len()])
 V2=np.zeros([y.len(),y.len()])
 
-Gam=np.zeros([y.len(),y.len()])
-
 V0=20
 
 iter1=0
-
-t=np.linspace(0,10,1001)
-tpot=potential(t,'fwell',np.array([4.,6.,V0]))
-
-plt.plot(t,tpot)
-plt.show()
-
 
 for e in y.e:
         b=e.matrix('d|d')
@@ -84,7 +75,7 @@ niter=15
 eps=1e-1j
 
 #Free Green's Operator
-n0=0
+n0=20
 nenergy=20
 store1=np.zeros([niter+2,nenergy])+0j
 
@@ -98,8 +89,9 @@ Tmat=Vmod
 
 
 for k in range(nenergy):
+	Gam=np.zeros([y.len(),y.len()])
 	for l in range(0,y.len()):
-		if evals[l]<0:	Gam=Gam+y.FEM_Outer(evecs[:,l],evecs[:,l])/(cos_evals[k+n0]-evals[l])
+		if evals[l]<0:	Gam=Gam+y.FEM_Outer(evecs[:,l],evecs[:,l])/(cos_evals[k+n0]+eps-evals[l])
 		else:	break
 
 	Emat=np.eye(y.len()) * (cos_evals[k+n0]+eps)
@@ -109,21 +101,26 @@ for k in range(nenergy):
 
 	G0=np.dot(Gtemp,y.overlap())
 	G_orig=np.dot(Gtemp2,y.overlap())
+
+#	Gam=np.dot(y.overlap(),Gam)
 	Gexact=G_orig-Gam
 
-	vec1=np.dot(Gexact,cos_evecs[:,k+n0])
-	store1[0,k]=np.dot(cos_evecs[:,k+n0],vec1)
+	#vec1=np.dot(Gexact,cos_evecs[:,k+n0])
+	#store1[0,k]=np.dot(cos_evecs[:,k+n0],vec1)
+	store1[0,k]=la.norm(Gexact)
 
-	vec1=np.dot(G_orig,cos_evecs[:,k+n0])
-	store1[1,k]=np.dot(cos_evecs[:,k+n0],vec1)
+#	vec1=np.dot(G_orig,cos_evecs[:,k+n0])
[…]

### 2012-05-13 python/test5plot/plotresults.py

@@ -2,8 +2,8 @@ import pickle as pickle
 import numpy as np
 import matplotlib.pyplot as plt
 
-f1=open('test5results3','r')
-f2=open('test5results3_unmodified','r')
+f1=open('test5results_new','r')
+f2=open('test5results_new_unmod','r')
 
 y1=pickle.load(f1)
 y2=pickle.load(f2)
@@ -12,7 +12,7 @@ y2=pickle.load(f2)
 plotval1=abs(y1[3:,:])
 plotval2=abs(y2[3:,:])
 
-n=np.linspace(0,14,15);
+n=np.linspace(0,19,20);
 
 #for k in range(4):
 #	mom_index=k+4
@@ -39,28 +39,27 @@ n=np.linspace(0,14,15);
 #
 #plt.show()
 
-for k in range(1):
-	mom_index=k+8
+for k in range(3):
+	mom_index=k
 
 	axlim1=.9*min(plotval1[:,mom_index])
 	axlim2=1.01*max(plotval1[:, mom_index])
-	yexact=abs(y1[0,mom_index]) * np.ones(15)
-	plt.subplot(1,2,k+1)
+	yexact=abs(y1[0,mom_index]) * np.ones(20)
+	plt.subplot(1,2,1)
 	plt.plot(n,plotval1[:,mom_index],'ro',n,yexact,'b--')
 	plt.plot(n,plotval1[:,mom_index],'r--')
 #	plt.axis([-1, 15, axlim1, axlim2])
 
 	axlim1=min(plotval2[:,mom_index])
 	axlim2=max(plotval2[:,mom_index])
-	yexact=abs(y2[0,mom_index]) * np.ones(15)
-	plt.subplot(1,2,k+2)
+	yexact=abs(y2[0,mom_index]) * np.ones(20)
+	plt.subplot(1,2,2)
 	plt.plot(n,plotval2[:,mom_index],'ro',n,yexact,'b--')
 	plt.plot(n,plotval2[:,mom_index],'r--')
 #	plt.axis([-1, 15, axlim1, axlim2])
-
-plt.suptitle('E_n=%s*pi^2/200'%(mom_index*mom_index))
-plt.figtext(.45,.02,'x-axis = no. of iterations')
-plt.show()
+	plt.suptitle('E_n=%s*pi^2/200'%(mom_index*mom_index))
+	plt.figtext(.45,.02,'x-axis = no. of iterations')
+	plt.show()
 
 	
 

### 2012-05-13 python/bornserieswtf.py
Sunday MAY 13th 2012 22:02

@@ -27,6 +27,7 @@ B=np.zeros([y.len(),y.len()])
 #Potential Matrix
 V1=np.zeros([y.len(),y.len()])
 V2=np.zeros([y.len(),y.len()])
+Gam_init=np.zeros([y.len(),y.len()])
 
 V0=20
 
@@ -68,8 +69,8 @@ for k in range(0,y.len()):
 	evecs[:,k]=evecs[:,k] / evecsnorm
 
 	#Potential Modification
-#	if evals[k]<0:
-#	        Gam=Gam + y.FEM_Outer(evecs[:,k],evecs[:,k])
+	if evals[k]<0:
+	        Gam_init=Gam_init + y.FEM_Outer(evecs[:,k],evecs[:,k])
 
 niter=15
 eps=1e-1j
@@ -77,7 +78,7 @@ eps=1e-1j
 #Free Green's Operator
 n0=20
 nenergy=20
-store1=np.zeros([niter+2,nenergy])+0j
+store1=np.zeros([niter+3,nenergy])+0j
 
 Bmod=np.dot(B/2.,y.overlap_inv())
 B_orig=np.dot(B/2. + V1 + V2,y.overlap_inv())
@@ -102,9 +103,17 @@ for k in range(nenergy):
 	G0=np.dot(Gtemp,y.overlap())
 	G_orig=np.dot(Gtemp2,y.overlap())
 
-#	Gam=np.dot(y.overlap(),Gam)
 	Gexact=G_orig-Gam
 
+#	A1=np.dot(np.dot(Gam_init,Gtemp2),Gam_init)
+#	A2=np.dot(Gam_init,Gtemp)
+#	A3=np.dot(Gtemp2,Gam_init)
+#	A4=np.dot(A3,A1)
+#	A5=np.dot(A4,A2)
+
+	A5=np.dot(Gam_init,Gtemp2)
+	print la.norm(Gam-A5)
+	raw_input()
 	#vec1=np.dot(Gexact,cos_evecs[:,k+n0])
 	#store1[0,k]=np.dot(cos_evecs[:,k+n0],vec1)
 	store1[0,k]=la.norm(Gexact)
@@ -113,14 +122,22 @@ for k in range(nenergy):
 #	store1[1,k]=np.dot(cos_evecs[:,k+n0],vec1)
 	store1[1,k]=la.norm(G_orig)
 
[…]

### 2012-07-26 Thesis - Presentation/draft1.tex
Thursday JUL 26th 2012 13:49

--- 2012-05-09 Thesis - Presentation/draft1.tex
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{graphicx}
-%\usepackage{abssymb}
+\usepackage{glossaries}
+\usepackage[menubordercolor={0 0 0}]{hyperref}
+\hypersetup{menubordercolor={0 0 0}}
 %opening
-\title{Thesis Draft 1}
+\title{Rapid Convergence of The Born Series}
 \author{Anupam Prasad Vedurmudi}
+\date{July 2012}
 
@@
 
-\section{Scattering Theory}
+\tableofcontents
+\newpage
+\chapter{Introduction}
+In this write-up we are concerned with the theoretical analysis of scattering experiments.
+The basic principle of scattering for classical particles is simple - we start with any number of particles 
+(two in the simplest case) which are far apart in some sense. This means that we can assume that the particles 
+don't interact with eachother in any way. After the passage of some time these particles ``collide'' or 
+interact locally in some sense and after some time they are far apart again. From the initial and final 
+properties (position, velocity, etc.) we can infer the details of the interaction.
 
-\subsection{Formal Scattering Theory}
-The spectrum of any
+In quantum scattering, however, matters are slightly more complicated. The very notion of a ``path'' with a definite velocity 
+is meaningless. We instead need to calculate the probability of a particle to deviate or ``scatter'' through a given angle. This
+quantity is known as the differential cross-section. Although it is the fundamental quantity of interest in scattering theory,
+our focus in this write-up will be on a tool that is indirectly related to it - the Born series. We note that quantum scattering
+theory can be further divided into nonrelativistic and relativistic cases. In this write-up we are only concerned with the former case.
 
+The Born series is a perturbation expansion that enables us to calculate the Scattering operator to a desired order. 
+The Scattering operator calculates the final state given a specific initial state. This can then be used to compute the
+cross-section. The objects of interest that can be studied using this method are the Green's function and the T operator.
+\chapter{Scattering Theory}
+
+This section is based on the textbook, \cite{TaylorScattering}
+and the first part of the lecture notes, \cite{YafaevScattering}. We will present the basic principles of quantum scattering
+and the use of the Born series. As an example, we use the problem of the scattering of a single particle off
+a general potential. It can be shown that this problem is equivalent to that of two-particle
+scattering.
 We define the Hamiltonian operator as,
@@
 \end{equation}
-\noindent Where, $H_0$ is the free Hamiltonian given by,
-
+ Where, $H_0$ is the free Hamiltonian given by, 
 \begin{equation}
@@
 \end{equation}
-\noindent $V$ is the potential.
+ $\Delta$ is the Laplace operator and $V$ corresponds to the potential. 
+For convenience, we work in units where, $\hbar^2/m=1$. Here, $m$ is the
+mass of the particle.
 
-\noindent The Green's function is formally defined as,
-\begin{eqnarray}\label{GFdefinition}
- G\left(E+i\epsilon\right) &=\left(E+i\epsilon-H\right)^{-1}\\
- G_0\left(E+i\epsilon\right) &=\left(E+i\epsilon-H_0\right)^{-1}
-\end{eqnarray}
+\section{Spectrum of The Hamiltonian}
+ $H_0$ and $H$ are self-adjoint operators on a Hilbert space
+$\mathcal{H}$. In general, the spectrum of a self-adjoint operator has
+two components,
+\begin{itemize}
+ \item Discrete - corresponding to the bound states
+ \item Continuous - corresponding to the scattering states
+\end{itemize}
+The continuous part of the spectrum can be further divided into into
+an $absolutely$ $continuous$ and a $singular\ continuous$ part. This results in a
+corresponding decomposition of the Hilbert space into three invariant subspaces.
+\begin{equation}\label{HilbertDecomp}
+ \mathcal{H}=\mathcal{H}^{pp}\oplus\mathcal{H}^{ac}\oplus\mathcal{H}^{sc}
+\end{equation}
+ Where, $\mathcal{H}^{pp}$ is the part of the Hilbert space corresponding to
+the discrete (pure point) part of the spectrum. For most systems of interest, the singular continous 
+part is absent. Scattering theory is essentially perturbation theory applied on the 
+absolutely continous part of the spectrum.
 
-\noindent The Lippmann-Schwinger equation for the Green's function is given by,
+The spectrum of $H_0$ coincides with its absolutely continuous part - the discrete and singular parts are empty. In principle, the 
+spectrum of $H_0$ is the entire complex plane, but we are only interested in the positive real axis including zero ($\mathbb{R}^+\cup\{0\}$).
+These correspond to free particles. We denote these eigenstates of
+$H_0$ in $3D$ as $|\mathbf{k}\rangle$, where $\mathbf{k}\in\mathbb{R}^3$. The eigenvalue corresponding to these eigenstates is $E=|\mathbf{k}|^2/2$,
+. These eigenstates are improper, i.e., they are not normalisable,
+\begin{equation}\label{FreeNormalisation}
+ \langle\mathbf{k}^{\prime}|\mathbf{k}\rangle=\delta\left(\mathbf{k}^{\prime}-\mathbf{k}\right)
+\end{equation}
+Where $\delta$ is the Dirac delta function. The reason for the notation is that, the wavefunction corresponding this eigenstate is given by,
+\begin{equation}\label{FreeWavefunction}
+  \langle\mathbf{x}|\mathbf{k}\rangle = e^{i\mathbf{k}.\mathbf{x}}
+\end{equation}
+viz., the plane-wave with wave vector $\mathbf{k}$.
+\subsubsection{Bound States}
+The eigenstates corresponding to the discrete part of the spectrum are called \emph{bound} states and those of the continuous
+part are called the \emph{scattering} states. The potentials we will be dealing with support at least one negative energy
+bound state. The divergence of the Born series at low energies is due to the presence of bound states close to $0$.
+We will later see that the scattering states asymptotically (in terms of time-evolution) 
+look like free-eigenstates while the bound states don't.
+\vspace{\baselineskip}
+
+\section{Time Evolution and The Asymptotic Condition}
+ In the Schrödinger picture, the time evolution of a state $|\psi_t\rangle$ is given by,
+\begin{equation}\label{SchroedingerEvolution1}
+ i\frac{d}{dt}|\psi_t\rangle = H|\psi_t\rangle
+\end{equation}
+Given an initial state $|\psi\rangle$, the formal solution of \eqref{SchroedingerEvolution1}
+given by,
+\begin{equation}\label{TimeEvolution}
+ |\psi_t\rangle = U(t)|\psi\rangle 
+\end{equation}
+Where, $U(t)=e^{-iHt}$. Since $H$ is self-adjoint, $U(t)$ is a unitary operator on $\mathcal{H}$, i.e., it is a linear operator that
+maps the whole of the Hilbert space onto itself and preserves the norm. Put more precisely, 
+\begin{equation*}
+\mathcal{D}\left(U(t)\right)=\mathcal{R}\left(U(t)\right)=\mathcal{H}
+\end{equation*}
+And, $\Arrowvert\Omega\psi\Arrowvert=\Arrowvert\psi\Arrowvert$. As a consequence, we are led a useful property of unitary operators, 
+\begin{equation}\label{unitaryprop}
+  U^\dagger(t)U(t)=U(t)U^\dagger(t)=\mathbb{I}
+\end{equation}
+
+The essence of scattering theory is that \eqref{TimeEvolution} describes the evolution of some scattering wavefunction such that,
+\begin{align}\label{Asymptotic1}
+ U(t)|\psi\rangle &\displaystyle\xrightarrow[t\rightarrow-\infty]{}  U^0(t)|\psi_{in}\rangle\\
+ U(t)|\psi\rangle &\displaystyle\xrightarrow[t\rightarrow+\infty]{}  U^0(t)|\psi_{out}\rangle
+\end{align}
+Where, $U^0(t)=e^{-iH_0t}$ is the free evolution operator and $|\psi_{in}\rangle$ and $|\psi_{out}\rangle$ are the in and out
+states. To be precise, the in and out states describe the behaviour of the system a long time before and after the interaction respectively.
+The in and out states behave like free wave-packets - i.e., their time evolution is governed by the free Hamiltonian.
+
+The asymptotic condition states that for every $|\psi_{in}\rangle$ in $\mathcal{H}$ there is a $|\psi\rangle$ such that 
+\eqref{Asymptotic1}.
+
+\subsection{Time-Independent Scattering Theory}
+\section{Green's Operator}
+The Green's operator is formally defined as,
+\begin{align}\label{GFdefinition}
+ G\left(E\pm i0\right) &=\left(E\pm i0-H\right)^{-1}\\
+ G_0\left(E\pm i0\right) &=\left(E\pm i0-H_0\right)^{-1}
+\end{align}
+Where, $G\left(E\pm i0\right)=\displaystyle\lim_{\epsilon\rightarrow 0}G(E\pm i\epsilon)$. The sign
+of $i\epsilon$ indicates the direction of time (positive for forward scattering). The
+Lippmann-Schwinger equation for the Green's operator is given by,
 \begin{equation}\label{LS1}
@@
 \end{equation}
+ The above equation can be derived by using the matrix identity,
+\begin{equation}\label{MatrixIdentity1}
+	A^{-1}=B^{-1}+A^{-1}\left(B-A\right)B^{-1}
+\end{equation}
 
-\noindent Subsequently, the Born series for $G$ is obtained by recursively substituting
+Subsequently, an operator-series expansion for $G$ is obtained by iteratively substituting
 the $(n-1)^{th}$ order approximation for $G$ in the R.H.S of \eqref{LS1} to get the $n^{th}$
-order term. The $0^{th}$ order approximation is the free Green's function $G_0$, the $1^{st}$
+order term. The $0^{th}$ order approximation is the free Green's operator $G_0$, the $1^{st}$
 order approximation (the Born approximation) is given by,
@@
 \end{equation}
-
-The Born series can thus be written as,
-
+The Born series for the Green's operator can thus be written as,
 \begin{equation}\label{BornSeriesExpanded}
@@
 \end{equation}
-
-\noindent Or more compactly,
+Or more compactly,
 \begin{equation}\label{BornSeriesCompact}
@@
 
-\section{Finite Element Method}
+\section{T-Operator}
+As described in \cite{TaylorScattering}, it is convenient to define the $T$-operator,
+\begin{equation}\label{Toperator1}
+ T=V+VGV
+\end{equation}
+ Multiplying both sides by $G_0$ we get and using \eqref{LS1}, we get,
+\begin{equation}\label{TOperatorRelation1}
+ G_0T=GV
+\end{equation}
+ Substituting this relation back in \eqref{Toperator1} we get a Lippmann-Schwinger equation
+for the $T$-Operator,
+\begin{equation}\label{LS2}
+ T=V+VG_0T
+\end{equation}
+ And correspondingly, through iteration, a Born Series,
+\begin{equation}\label{BornSeriesToperator}
+ T=\displaystyle\sum_{n=0}^{\infty}V\left(G_0V\right)^n
+\end{equation}
+The Born approximation in this case is, $T\simeq V$.
+
+The momentum space matrix elements of the T-matrix are given by,
+\begin{equation}\label{Tmatrixelements}
+  t\left(\mathbf{k}^\prime\leftarrow\mathbf{k}\right)=\langle \mathbf{k}^\prime |T| \mathbf{k} \rangle
+\end{equation}
+
+\section{Scattering Operator}
+We first define the Møller operators, $\Omega_\pm$ as,
+\begin{equation}\label{moellerdef}
+ \Omega_\pm = \displaystyle\lim_{t\rightarrow\mp\infty}U(t)^\dagger U^0(t)
+\end{equation}
+These operators act on the in and out states to give,
+\begin{align}\label{moelleract}
+ |\psi\rangle &=\Omega_+|\psi_{in}\rangle \\
+ |\psi\rangle &=\Omega_-|\psi_{out}\rangle 
+\end{align}
+
+The Møller wave operators acting on an in or out state give the corresponding scattering state
+for the Hamiltonian. These operators are isometric, i.e., they are defined on all of $\mathcal{H}$
+and preserve the norm. They differ from unitary operators in that, $\mathcal{R}\left(\Omega_\pm\right)\neq\mathcal{H}$.
+Unlike the property \eqref{unitaryprop} for unitary operators, we have $\Omega_\pm^\dagger\Omega_\pm=\mathbb{I}$, but
+in general, $\Omega_\pm\Omega^\dagger_\pm\neq\mathbb{I}$. We also note that all Unitary operators are isometric.We 
+can now define the scattering operator,
+\begin{equation}\label{scatteringopdef}
+ \mathcal{S}=\Omega_-^\dagger\Omega_+
+\end{equation}
+The scattering operator has the property that,
+\begin{equation}\label{scatterinprop}
+ |\psi_{out}\rangle = \mathcal{S}|\psi_{in}\rangle
+\end{equation}
+i.e., given an in state, the scattering operator gives the corresponding out state. An important property of the $\mathcal{S}$ operator is
+that, unlike the Møller operators, it is unitary. As a consequence of this, every $|\psi_{in}\rangle$ is mapped to a unique $|\psi_{out}\rangle$ in $\mathcal{H}$.
+
+In practice, only the asymptotic free motion 
+is observable. Therefore, the operator $\mathcal{S}$ contains all the information of experimental interest. For a scattering process,
+$|\psi_{in}\rangle\rightarrow|\psi_{out}\rangle$, the probability is given by,
+\begin{equation}\label{scatteringprob}
+ \mathbf{w}(|\psi_{out}\rangle\leftarrow|\psi_{in}\rangle)=\lvert\langle\psi_{out}\rvert \mathcal{S}\lvert\psi_{in}\rangle\rvert^2
+\end{equation}
+
+The momentum space matrix elements of the scattering operator are related to \eqref{Tmatrixelements} as,
+\begin{equation}\label{Smatrixelements}
+ \langle \mathbf{k}^\prime |\mathcal{S}| \mathbf{k} \rangle = \delta_3(\mathbf{k}^\prime - \mathbf{k})-2\pi i\delta\left(E_{p^\prime}-E_p\right)t\left(\mathbf{k}^\prime\leftarrow\mathbf{k}\right)
+\end{equation}
+The physical interpretation of the above expression is that, given an incoming plane wave, after the interaction the outgoing wave
+is given by the incoming wave plus a component which a consequence of the interaction. The $T$ operator in a sense contains all
+the information about the interaction.
+Thus we see that, given a good approximation to the Green's operator or the $T$-operator, we can directly calculate the $S$ operator
+and thereby the cross-sections.
+\chapter{Finite Element Method}
 The Finite Element Method (FEM) is a numerical technique used to solve differential equations.
@@
 \begin{equation}\label{overlap}
- S_{nm}=\int d^3x\phi_n(\mathbf{x})\phi_m(\mathbf{x})
+ S_{nm}=\int J(\mathbf{x})d^nx\phi_n(\mathbf{x})\phi_m(\mathbf{x})
 \end{equation}
-
 We also define the matrix,
 \begin{equation}\label{diff_overlap}
- B_{nm}=\int d^3x\nabla\phi_n(\mathbf{x}).\nabla\phi_m(\mathbf{x})
+ B_{nm}=\int J(\mathbf{x})d^nx\nabla\phi_n(\mathbf{x}).\nabla\phi_m(\mathbf{x})
 \end{equation}
-
-\subsection{Boundary Conditions}
+Here, $J(\mathbf{x})$ is the Jacobian for the chosen coordinate system.
+\section{Boundary Conditions}
 In order to solve the differential equations we must also specify appropriate boundary
@@
 
-\subsubsection{Dirichlet Boundary Condition}
+\subsection{Dirichlet Boundary Condition}
 The Dirichlet boundary condition requires that the function has a known value on the boundary.
@@
 
-\subsubsection{Neumann Boundary Condition}
+\subsection{Neumann Boundary Condition}
 The Neumann boundary condition requires that the gradient of the function
@@
 
-\subsubsection{Mixed Boundary Condition}
+\subsection{Mixed Boundary Condition}
 It is also possible to have a boundary such that the function has a Dirichlet boundary
@@
 
-\subsection{Python Implementation}
-\section{FEM Formulation of Quantum Mechanics}
+\subsubsection{Continuity}
+The function is further constrained to be continuous at the boundaries of the finite element.
+
+\subsubsection{Note}
+We will be working in 1-dimension with a Jacobian, $J(x)=1$. We should also mention here that, a $2D$ problem in polar coordinates with azimuthal symmetry reduces
+to a $1D$ problem with a Jacobian, $J(x)=x$ and similarly, a $3D$ problem in spherical coordinates with symmetry in both angular direction reduces to
+a $1D$ problem with a Jacobian, $J(x)=x^2$. The python implementation is discussed in Appendix \ref{pythonimp}.
+
+\chapter{FEM in Quantum Mechanics}
 The Schrödinger equation for a wave function $\psi$ is given by,
-
 \begin{equation}\label{schroedinger1}
@@
 \end{equation}
-
-\noindent Using the FEM approximation defined in \eqref{femfirst},
+Using the FEM approximation defined in \eqref{femfirst},
 \begin{equation}\label{schroedinger}
@@
 \end{equation}
-\noindent We get the weak formulation of the problem by multiplying both sides by $\phi_m(\mathbf{x})$ and integrating over. After integrating by parts, the kinetic term gives us
-the following matrix element,
+
+Now, we get the weak formulation of the problem by multiplying both sides by $\phi_m(\mathbf{x})$ and integrating over. We now integrate by parts and use the finite
+support of the basis functions to set the boundary terms to zero. We have to note here that the integration is over all space not just the "box" of our problem. Even
+if the wavefunction has a non-zero value on the boundary of the box, the finite element basis functions go to zero as $|\mathbf{x}|\rightarrow\infty$
+
+Now we use the definition in \eqref{diff_overlap}, to get the following matrix element,
 \begin{equation}\label{kineticterm}
- \int d^3x\phi_m(\mathbf{x})\Delta\phi_n(\mathbf{x}) = -\int d^3x\nabla\phi_m(\mathbf{x}).\nabla\phi_n(\mathbf{x}) = -B_{mn}
+ B_{mn} = -\int d^3x\phi_m(\mathbf{x})\Delta\phi_n(\mathbf{x}) = \int d^3x\nabla\phi_m(\mathbf{x}).\nabla\phi_n(\mathbf{x})
+\end{equation}
+This is the FEM representation of the kinetic term.
+The potential term is calculated from the known functions $\phi_n$ and is defined as,
+\begin{equation}\label{potentialterm}
+  V_{mn} = \int d^3xV(\mathbf{x})\phi_m(\mathbf{x})\phi_n(\mathbf{x})
+\end{equation}
+Finally, using the definition of the overlap matrix in \eqref{overlap}, the R.H.S is given by,
+\begin{equation}\label{qmoverlap}
+ ES_{mn}=E\int d^3x\phi_m(\mathbf{x})\phi_n(\mathbf{x})
 \end{equation}
 
-\noindent The potential term is calculated from the known functions $\phi_n$ and is defined as,
-\begin{equation}\label{potentialterm}
- \int d^3xV(\mathbf{x})\phi_m(\mathbf{x})\phi_n(\mathbf{x}) = V_{mn}
-\end{equation}
-
-\noindent Finally, the R.H.S is given by,
-\begin{equation}
- E\int d^3x\phi_m(\mathbf{x})\phi_n(\mathbf{x})=ES_{mn}
-\end{equation}
-
-\noindent We define the coefficient vector $\mathbf{\underline{c}}=[c_1,c_2,\ldots,c_N]^T$. The Schrödinger equation reduces to 
-the following eigenvalue problem,
+Finally, we define the coefficient vector, $\mathbf{\underline{c}}=[c_1,c_2,\ldots,c_N]^T$. The Schrödinger equation thus reduces to 
+the following generalized eigenvalue problem,
 \begin{equation}\label{schroedingerfem}
@@
 \end{equation}
-\noindent Where, $H=\left(\frac{B}{2}+V\right)$
+Where, $H=\left(\frac{B}{2}+V\right)$
 
+\subsubsection{Inner Products}
 The inner product of two wavefunctions $\phi$ and $\chi$ is given by,
-\begin{eqnarray}\label{innerproduct}
- \int d^3x \chi^*(\mathbf{x})\psi(\mathbf{x}) &=& \displaystyle\sum_{m,n}^{N}c_md_n\int d^3x \phi_m(\mathbf{x})\phi_n(\mathbf{x})\\
- &=& \mathbf{c}^{\dagger}S\mathbf{d}
-\end{eqnarray}
-\noindent Where, $\mathbf{c}^{\dagger}=(\mathbf{c}^*)^T$.
+\begin{align}\label{innerproduct}
+ \int d^3x \chi^*(\mathbf{x})\psi(\mathbf{x}) &= \displaystyle\sum_{m,n}^{N}c^*_md_n\int d^3x \phi_m(\mathbf{x})\phi_n(\mathbf{x}) \nonumber\\
+ &= \mathbf{c}^{\dagger}S\mathbf{d}
+\end{align}
+ Where, $\mathbf{c}^{\dagger}=(\mathbf{c}^*)^T$ and we used the definition of the overlap matrix in \eqref{qmoverlap}.
 
-\subsection{Momentum Eigenstates}
+\section{Momentum Eigenstates}
 We now want to find an FEM representation for free eigenstate of momentum $\mathbf{k}$ (energy,
-$E_k=|\mathbf{k}|^2/2$). 
+$E_k=|\mathbf{k}|^2/2$). These are the improper eigenstates of the operator $H_0=B/2$. 
 It is most convenient to do so by defining a ``cosine eigenvector'',
 \begin{equation}\label{cosevec}
- |\mathbf{k}\rangle = exp(-i\mathbf{k}.\mathbf{x}) + exp(i\mathbf{k}.\mathbf{x})
+ |\mathbf{k}_c\rangle = exp(-i\mathbf{k}.\mathbf{x}) + exp(i\mathbf{k}.\mathbf{x})
 \end{equation}
 These can be obtained by ``solving'' the free Schrödinger equation on a grid with open
-boundaries, i.e., by solving the eigenvalue problem,
+boundaries, i.e., by solving \eqref{schroedingerfem} with $H=H_0$, i.e.,
 
 \begin{equation}\label{freeschroedinger}
- H_0\psi=\lambda S\psi
+ H_0\psi=E_k S\psi
 \end{equation}
@@
 
-$|\mathbf{k}\rangle$ has the following properties,
-\begin{eqnarray}
- \langle\mathbf{k}|S|\mathbf{k}\rangle &=&\int_{\Omega} d^3x\\
- \langle\mathbf{k}^{\prime}|S|\mathbf{k}\rangle &\simeq& 0
-\end{eqnarray}
-Here, $\int_{\Omega} d^3x$ is the volume of the solution space.
+$|\mathbf{k}_c\rangle$ has the following properties,
+\begin{align}
+ \langle\mathbf{k}_c|S|\mathbf{k}_c\rangle &= \int_{\Omega} d^3x = V_\Omega\\
+ \langle\mathbf{k}_c^{\prime}|S|\mathbf{k}_c\rangle &\simeq 0
+\end{align}
+Here, $V_\Omega$ is the volume of the solution space.
 
-\subsection{Born Series - FEM Formulation}
-We calculate the FEM representation of the Greens Function using \eqref{schroedingerfem}. We require that,
+\section{Born Series - FEM Formulation}
+We now use the definition of the Greens function in \eqref{GFdefinition} and the FEM formulation of the
+Schrödinger equation, \eqref{schroedingerfem} to calculate the FEM representation of the Greens Function. 
+We require that,
 \begin{equation}\label{greenfemreq}
@@
 \end{equation}
-\noindent Where, $\psi$ is an eigenstate of $H$ of energy $E_0$. Thus, we define the Green's function as,
+ Where, $\psi$ is an eigenstate of $H$ of energy $E_0$ and $z\neq E_0$ is a complex number. Thus, we define the Green's operator as,
 \begin{equation}\label{greenfemdef}
@@
 \end{equation}
-\noindent Where, $\mathbf{I}$ is the identity matrix. The free Greens function is defined similarly by setting
-$H=B/2$. We will ultimately be testing the convergence of, $\langle\mathbf{k}|G_{fem}(z)|\mathbf{k}\rangle$.
+ Where, $\mathbf{I}$ is the identity matrix. The free Greens function is defined similarly by setting
+$H=B/2$. We will ultimately be testing the convergence of the matrix element, $\langle\mathbf{k}|G_{fem}(z)|\mathbf{k}\rangle$.
 For later convenience, the potential matrix is also suitably modified as,
@@
 
-\subsection{Modification of the Hamiltonian}
-In this section we use the techniques outlined in \cite{KukPom76}. It this paper, the Born
-series is rearranged by using orthogonally projecting pseudopotentials (OPP). It is then
-proved that the rearranged Born series converges for all negative and small positive energies
-even if the system contains bound states.
+\chapter{Orthogonal Projection Method}
+In this section we use the techniques introduced in \cite{KukPom76} and elaborated in
+\cite{KukPom78}. In this paper, it is suggested that the Born series can be rearranged 
+by using orthogonally projecting pseudopotentials (OPP). It is then proved that the 
+rearranged Born series converges for all negative and small positive energies even if 
+the system contains bound states.
 
@@
 \end{equation}
-\noindent The rate of convergence of the Born series is controlled by the parameter $\lambda$.
+The rate of convergence of the Born series is controlled by the parameter $\lambda$. It is
+seen that for weak potentials, $\lambda$ has very little influence on the convergence while
+for stronger potentials which support several bound states, $\lambda$ seems to have
+an optimal range for convergence.
 
 The modified Hamiltonian and the modified free Hamiltonian are denoted by,
-
-\begin{eqnarray}
- \widetilde{H}&=&H+\Gamma\\
- \widetilde{H}_0&=&H_0+\Gamma
-\end{eqnarray}
-
+\begin{align}
+ \widetilde{H}&=H+\Gamma\\
+ \widetilde{H}_0&=H_0+\Gamma
+\end{align}
 The modified Greens functions $\widetilde{G}$ and $\widetilde{G}_0$ are defined similarly as
@@
 \end{equation}
-
-\noindent Physically, this means that the modified Hamiltonian has the same continuous spectrum but the discrete
+Physically, this means that the modified Hamiltonian has the same continuous spectrum but the discrete
 spectrum has been shifted by $\lambda$.
 
-\section{Toy Problem - Scattering in 1D}
+\section{Example Problem - Scattering in 1D}
 We now use the above methods in an example with scattering off a potential in 1D. Due to the simplicity of the problem,
-it is possible to explicitly calculate the exact Green's function by inverting the appropriate matrix according the formal
+it is possible to explicitly calculate the exact Green's operator by inverting the appropriate matrix according the formal
 definition given in \eqref{GFdefinition}. We will examine the problem in two cases: 
@@
 
-Our box is defined as an FEM axis from $x=0$ to $x=10$ with $n=400$ discretization points, 
-of order $40$ and with open boundaries. This means that the axis is divided into 400 
-parts and each part has functions (elements) upto order $40$. Order $1$ being a straight
-line. As an example, we show an axis, $0\leq x\leq 4$ with $n=4$ discretization points, open
+Our box is defined as an FEM axis from $x=0$ to $x=10$ with $N=200$ discretization points, 
+of order $40$ and with open boundaries. This means that the axis is divided into 200 
+segments and each segment has functions (elements) upto order $40$. Order $1$ being a straight
+line. As an example, we show an axis, $0\leq x\leq 4$ with $N=4$ discretization points, open
 boundaries and order $1$.
@@
 \includegraphics[width=225pt, height=150pt]{femexample1.png}
-\caption[\textwidth]{FEM axis. n=4, order=1}
+\caption[\textwidth]{FEM axis. N=4, order=1}
 \end{figure}
-The different elements are distinguished by colour.
 
 The ``cosine eigenvectors'' are computed using \eqref{cosevec} and \eqref{freeschroedinger}.
-The energies of the free wavefunctions are $E_n = n^2\pi^2/200$ for $n=0,\ldots,19$. The factor
-of $200$ is a result of the discretization. The Green's function, $G(z)$ is calculated for $z=E_n+i\epsilon$,
-where, $\epsilon$ is a small positive number defined based on the problem. The purpose of epsilon is to avoid 
-the singularity on the positive real axis.
+The energies of the free wavefunctions are $E_n = n^2\pi^2/2L^2$ for $n=0,\ldots,N$ and $L=10$.
+The Green's operator, $G(z)$ is calculated for $z=E_n+i\epsilon$,
+where $\epsilon$ is a small positive number. The purpose of $\epsilon$ is to avoid 
+the singularity of the Green's operator on the positive real axis. We first encountered it in \eqref{GFdefinition} where
+we originally took $\epsilon\rightarrow 0$. In the numerical example, however, epsilon is fixed to be a small
+number which is larger than the energy spacing between the bound states close to the origin.
 
 \subsection{First Case}
-In the first case, we define the potential as a finite well of strength $0.1$ with support in $4 \leq x \leq 6$,
+%\begin{center}
+\begin{figure}
+\includegraphics[width=320pt, height=145pt]{series1a.png}
+\includegraphics[width=320pt, height=145pt]{series1b.png}
+\includegraphics[width=320pt, height=145pt]{series1c.png}
+\caption[\textwidth]{First Case - Weak Finite Well}
+\end{figure}
+%\end{center}
+
+In the first case, we define the potential as a finite well of strength $0.01$ with support in $4.5 \leq x \leq 5.5$,
 \begin{equation}\label{potentialdef1}
 V = \begin{cases}
--0.1 &\mbox{} 4\leq x\leq 6\\
+-0.01 &\mbox{} 4.5\leq x\leq 5.5\\
 0 &\mbox{ otherwise}
@@
 \end{equation}
+The Born Series is calculated for $15$ iterations. The Hamiltonian is modified using the operator defined in \eqref{potmod} with
+$\lambda=1$.  We also set $\epsilon=10^{-3}$.
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=225pt, height=150pt]{potential1d1.png}
-\caption[\textwidth]{Test Potential 1}
-% 
-% \centering
-% \includegraphics[width=225pt, height=150pt]{potential1d2.png}
-% \caption[\textwidth]{Test Potential 2}
-\end{figure}
-
-The Born Series is calculated for $15$ iterations. The Hamiltonian is modified using the operator defined in \eqref{potmod} with
-$\lambda=1$.  We also set $\epsilon=0.1$.
-
-We see that the Born Series converges to the exact value in both cases but that the modified case converges in fewer iterations.
+We now plot the elements of the Born Series against the number of iterations. In order
+to compare the modified and unmodified cases, we plot them side by side with the modified case on the left and the unmodified one
+on the right. We follow this convention in the other next case as well.
 
 \subsection{Second Case}
-In the second case, we define the potential as, 
-\begin{equation}\label{potentialdef2}
-V = -200\mathtt{exp}(-(x-3.5)^2/0.5)-200\mathtt{exp}(-(x-6.5)^2/0.5)
+The potential is give by,
+\begin{equation}\label{potentialdef3}
+V = -100\mathtt{exp}(-(x-3.5)^2/0.5)-100\mathtt{exp}(-(x-6.5)^2/0.5)
 \end{equation}
-
+We set $\epsilon=10^{-6}$ and $\lambda=100$.
 \begin{figure}[hb]
@@
 \end{figure}
+\begin{figure}
+\includegraphics[width=320pt, height=145pt]{series3a.png}
+\end{figure}
+\begin{figure}[ht]
+%\includegraphics[width=320pt, height=145pt]{series3a.png}
+\includegraphics[width=320pt, height=145pt]{series3b.png}
+\includegraphics[width=320pt, height=145pt]{series3c.png}
+\caption[\textwidth]{Gaussian Like Potential}
+\end{figure}
 
-The born series is calculated for $15$ iterations. We choose $\lambda=700$ and $\epsilon=10^{-8}$.
+We see that for some energies the Born Series in the modified case converges to the exact value but diverges in the original case. We also note that
+for certain energies, the Born Series diverges in both cases but the divergence in the unmodified case is slower.
 
-We see that the Born Series in the modified case converges to the exact value but diverges in the original case.
+\chapter{Conclusion}
+
+\appendix
+\addcontentsline{toc}{chapter}{Appendices}
+%\chapter{Calculation of The Scattering Cross-section}\label{crosssection}
+\chapter{Python Implementation}\label{pythonimp}
+The main object of interest is the \verb=Axis= class, which is defined in the file \verb=axis.py=.
 

### 2012-07-26 Thesis - Presentation/draft2.tex
Thursday JUL 26th 2012 13:49

--- 2012-05-11 Thesis - Presentation/draft1.tex
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{graphicx}
-%\usepackage{abssymb}
+\usepackage{glossaries}
+\usepackage[menubordercolor={0 0 0}]{hyperref}
+\hypersetup{menubordercolor={0 0 0}}
 %opening
-\title{Draft 1}
+\title{Theoretikum}
 \author{Anupam Prasad Vedurmudi}
@@
 
-\section{Scattering Theory}
+\tableofcontents
+\newpage
+\section{Introduction}
+In this write-up we are concerned with the theoretical analysis of scattering experiments.
+The basic principle of scattering for classical particles is simple - we start with any number of particles 
+(two in the simplest case) which are far apart in some sense. This means that we can assume that the particles 
+don't interact with eachother in any way. After the passage of some time these particles ``collide'' or 
+interact locally in some sense and after some time they are far apart again. From the initial and final 
+properties (position, velocity, etc.) we can infer the details of the interaction.
 
-\subsection{Formal Scattering Theory}
-The spectrum of any
+In quantum scattering, however, matters are slightly more complicated. The very notion of a ``path'' with a definite velocity 
+is meaningless. We instead need to calculate the probability of a particle to deviate or ``scatter'' through a given angle. This
+quantity is known as the differential cross-section. Although it is the fundamental quantity of interest in scattering theory,
+our focus in this write-up will be on a tool that is indirectly related to it - the Born series. We note that quantum scattering
+theory can be further divided into nonrelativistic and relativistic cases. In this write-up we are only concerned with the former case.
 
+The Born series is a perturbation expansion that enables us to calculate the Scattering operator to a desired order. 
+The Scattering operator calculates the final state given a specific initial state. This can then be used to compute the
+cross-section. The objects of interest that can be studied using this method are the Green's function and the T operator.
+\section{Preliminaries}
+
+This section is based on the textbook, \cite{TaylorScattering}
+and the first part of the lecture notes, \cite{YafaevScattering}. The problem is of the scattering of a single particle off
+a potential. It can be shown that this problem is equivalent to that of two-particle
+scattering.
 We define the Hamiltonian operator as,
@@
 \end{equation}
-\noindent $V$ is the potential.
+ $\Delta$ is the Laplace operator and $V$ corresponds to the potential. 
+For convenience, we work in units where, $\hbar^2/m=1$. Here, $m$ is the
+mass of the particle.
 
-\noindent The Green's function is formally defined as,
-\begin{eqnarray}\label{GFdefinition}
- G\left(E+i\epsilon\right) &=\left(E+i\epsilon-H\right)^{-1}\\
- G_0\left(E+i\epsilon\right) &=\left(E+i\epsilon-H_0\right)^{-1}
-\end{eqnarray}
+\subsection{Spectrum of The Hamiltonian}
+ $H_0$ and $H$ are self-adjoint operators on a Hilbert space
+$\mathcal{H}$. In general, the spectrum of a self-adjoint operator has
+two components,
+\begin{itemize}
+ \item Discrete - corresponding to the bound states
+ \item Continuous - corresponding to the scattering states
+\end{itemize}
+The continuous part of the spectrum can be further divided into into
+an $absolutely$ $continuous$ and a $singular\ continuous$ part. This results in a
+corresponding decomposition of the Hilbert space into three invariant subspaces.
+\begin{equation}\label{HilbertDecomp}
+ \mathcal{H}=\mathcal{H}^{pp}\oplus\mathcal{H}^{ac}\oplus\mathcal{H}^{sc}
+\end{equation}
+ Where, $\mathcal{H}^{pp}$ is the part of the Hilbert space corresponding to
+the discrete (pure point) part of the spectrum. For most systems of interest, the singular continous 
+part is absent. Scattering theory is essentially perturbation theory applied on the 
+absolutely continous part of the spectrum.
 
-\noindent The Lippmann-Schwinger equation for the Green's function is given by,
+The spectrum of $H_0$ only contains an absolutely continuous part - the positive real axis including $0$. We denote the eigenstates of
+$H_0$ as $|\mathbf{k}\rangle$, where $\mathbf{k}\in\mathbb{R}^3$. The eigenvalue corresponding to these eigenstates is $E=k^2/2$,
+$k=|\mathbf{k}|$. These eigenstates are improper, i.e., they are not normalisable,
+\begin{equation}\label{FreeNormalisation}
+ \langle\mathbf{k}^{\prime}|\mathbf{k}\rangle=\delta\left(\mathbf{k}^{\prime}-\mathbf{k}\right)
+\end{equation}
+Where $\delta$ is the Dirac delta function. The reason for the notation is that, the wavefunction corresponding this eigenstate is given by,
+\begin{equation}\label{FreeWavefunction}
+  \langle\mathbf{x}|\mathbf{k}\rangle = e^{i\mathbf{k}.\mathbf{x}}
+\end{equation}
+The eigenstates corresponding to the discrete part of the spectrum are called the \emph{bound} states and those of the continuous
+part are called the \emph{scattering} states. We will later see that the scattering states asymptotically look like free-eigenstates while
+the bound states don't.
+\vspace{\baselineskip}
+
+\subsection{Time Evolution and The Asymptotic Condition}
+ In the Schrödinger picture, the time evolution of a state $|\psi_t\rangle$ is given by,
+\begin{equation}\label{SchroedingerEvolution1}
+ i\frac{d}{dt}|\psi_t\rangle = H|\psi_t\rangle
+\end{equation}
+Given an initial state $|\psi\rangle$, the formal solution of \eqref{SchroedingerEvolution1}
+given by,
+\begin{equation}\label{TimeEvolution}
+ |\psi_t\rangle = U(t)|\psi\rangle 
+\end{equation}
+Where, $U(t)=e^{-iHt}$. Since $H$ is self-adjoint, $U(t)$ is a unitary operator on $\mathcal{H}$, i.e., it is a linear operator that
+maps the whole of the Hilbert space onto itself and preserves the norm. Put more precisely, 
+\begin{equation*}
+\mathcal{D}\left(U(t)\right)=\mathcal{R}\left(U(t)\right)=\mathcal{H}
+\end{equation*}
+And, $\Arrowvert\Omega\psi\Arrowvert=\Arrowvert\psi\Arrowvert$. As a consequence, we are led a useful property of unitary operators, 
+\begin{equation}\label{unitaryprop}
+  U^\dagger(t)U(t)=U(t)U^\dagger(t)=\mathbb{I}
+\end{equation}
+
+The essence of scattering theory is that \eqref{TimeEvolution} describes the evolution of some scattering wavefunction such that,
+\begin{align}\label{Asymptotic1}
+ U(t)|\psi\rangle &\displaystyle\xrightarrow[t\rightarrow-\infty]{}  U^0(t)|\psi_{in}\rangle\\
+ U(t)|\psi\rangle &\displaystyle\xrightarrow[t\rightarrow+\infty]{}  U^0(t)|\psi_{out}\rangle
+\end{align}
+Where, $U^0(t)=e^{-iH_0t}$ is the free evolution operator and $|\psi_{in}\rangle$ and $|\psi_{out}\rangle$ are the in and out
+states. The in and out states behave like free wave-packets - i.e., their time evolution is governed by the free Hamiltonian.
+
+The asymptotic condition states that for every $|\psi_{in}\rangle$ in $\mathcal{H}$ there is a $|\psi\rangle$ such that 
+\eqref{Asymptotic1}.
+
+\subsection{Green's Operator}
+The Green's operator is formally defined as,
+\begin{align}\label{GFdefinition}
+ G\left(E\pm i0\right) &=\left(E\pm i0-H\right)^{-1}\\
+ G_0\left(E\pm i0\right) &=\left(E\pm i0-H_0\right)^{-1}
+\end{align}
+Where, $G\left(E\pm i0\right)=\displaystyle\lim_{\epsilon\rightarrow 0}G(E\pm i\epsilon)$. The
+Lippmann-Schwinger equation for the Green's operator is given by,
 \begin{equation}\label{LS1}
@@
 \end{equation}
-
-\noindent Or more compactly,
+Or more compactly,
 \begin{equation}\label{BornSeriesCompact}
  G=\displaystyle\sum_{n=0}^{\infty}G_0\left(VG_0\right)^n
+\end{equation}
+
+\subsection{T-Operator}
+As described in \cite{TaylorScattering}, it is convenient to define the $T$-operator,
+\begin{equation}\label{Toperator1}
+ T=V+VGV
+\end{equation}
+ Multiplying both sides by $G_0$ we get and using \eqref{LS1}, we get,
+\begin{equation}\label{TOperatorRelation1}
+ G_0T=GV
+\end{equation}
+ Substituting this relation back in \eqref{Toperator1} we get a Lippmann-Schwinger equation
+for the $T$-Operator,
+\begin{equation}\label{LS2}
+ T=V+VG_0T
+\end{equation}
+ And correspondingly, through iteration, a Born Series,
+\begin{equation}\label{BornSeriesToperator}
+ T=\displaystyle\sum_{n=0}^{\infty}V\left(G_0V\right)^n
+\end{equation}
+The Born approximation in this case is, $T\simeq V$.
+
+The momentum space matrix elements of the T-matrix are given by,
+\begin{equation}\label{Tmatrixelements}
+  t\left(\mathbf{k}^\prime\leftarrow\mathbf{k}\right)=\langle \mathbf{k}^\prime |T| \mathbf{k} \rangle
+\end{equation}
+
+\subsection{Scattering Operator}
+We first define the Møller operators, $\Omega_\pm$ as,
+\begin{equation}\label{moellerdef}
+ \Omega_\pm = \displaystyle\lim_{t\rightarrow\mp\infty}U(t)^\dagger U^0(t)
+\end{equation}
+These operators act on the in and out states to give,
+\begin{align}\label{moelleract}
+ |\psi\rangle &=\Omega_+|\psi_{in}\rangle \\
+ |\psi\rangle &=\Omega_-|\psi_{out}\rangle 
+\end{align}
+
+The Møller wave operators acting on an in or out state give the corresponding scattering state
+for the Hamiltonian. These operators are isometric, i.e., they are defined on all of $\mathcal{H}$
+and preserve the norm. They differ from unitary operators in that, $\mathcal{R}\left(\Omega_\pm\right)\neq\mathcal{H}$.
+Unlike the property \eqref{unitaryprop} for unitary operators, we have $\Omega_\pm^\dagger\Omega_\pm=\mathbb{I}$, but
+in general, $\Omega_\pm\Omega^\dagger_\pm\neq\mathbb{I}$. We also note that all Unitary operators are isometric.We 
+can now define the scattering operator,
+\begin{equation}\label{scatteringopdef}
+ S=\Omega_-^\dagger\Omega_+
+\end{equation}
+The scattering operator has the property that,
+\begin{equation}\label{scatterinprop}
+ |\psi_{out}\rangle = S|\psi_{in}\rangle
+\end{equation}
+i.e., given an in state, the scattering operator gives the corresponding out state. In practice, only the asymptotic free motion 
+is observable. Therefore, the operator $S$ contains all the information of experimental interest. For a scattering process,
+$|\psi_{in}\rangle\rightarrow|\psi_{out}\rangle$, the probability is given by,
+\begin{equation}\label{scatteringprob}
+ \mathbf{w}(|\psi_{out}\rangle\leftarrow|\psi_{in}\rangle)=\lvert\langle\psi_{out}\rvert S\lvert\psi_{in}\rangle\rvert^2
+\end{equation}
+
+The momentum space matrix elements of the scattering operator are related to \eqref{Tmatrixelements} as,
+\begin{equation}\label{Smatrixelements}
+ \langle \mathbf{k}^\prime |S| \mathbf{k} \rangle = \delta_3(\mathbf{k}^\prime - \mathbf{k})-2\pi i\delta\left(E_{p^\prime}-E_p\right)t\left(\mathbf{k}^\prime\leftarrow\mathbf{k}\right)
 \end{equation}
@@
 
-\noindent Using the FEM approximation defined in \eqref{femfirst},
+ Using the FEM approximation defined in \eqref{femfirst},
 \begin{equation}\label{schroedinger}
@@
 \end{equation}
-\noindent We get the weak formulation of the problem by multiplying both sides by $\phi_m(\mathbf{x})$ and integrating over. After integrating by parts, the kinetic term gives us
+ We get the weak formulation of the problem by multiplying both sides by $\phi_m(\mathbf{x})$ and integrating over. After integrating by parts, the kinetic term gives us
 the following matrix element,
@@
 
-\noindent The potential term is calculated from the known functions $\phi_n$ and is defined as,
+ The potential term is calculated from the known functions $\phi_n$ and is defined as,
 \begin{equation}\label{potentialterm}
@@
 
-\noindent Finally, the R.H.S is given by,
+ Finally, the R.H.S is given by,
 \begin{equation}
@@
 
-\noindent We define the coefficient vector $\mathbf{\underline{c}}=[c_1,c_2,\ldots,c_N]^T$. The Schrödinger equation reduces to 
+ We define the coefficient vector $\mathbf{\underline{c}}=[c_1,c_2,\ldots,c_N]^T$. The Schrödinger equation reduces to 
 the following eigenvalue problem,
@@
 \end{equation}
-\noindent Where, $H=\left(\frac{B}{2}+V\right)$
+ Where, $H=\left(\frac{B}{2}+V\right)$
 
 The inner product of two wavefunctions $\phi$ and $\chi$ is given by,
-\begin{eqnarray}\label{innerproduct}
+\begin{align}\label{innerproduct}
  \int d^3x \chi^*(\mathbf{x})\psi(\mathbf{x}) &=& \displaystyle\sum_{m,n}^{N}c_md_n\int d^3x \phi_m(\mathbf{x})\phi_n(\mathbf{x})\\
  &=& \mathbf{c}^{\dagger}S\mathbf{d}
-\end{eqnarray}
-\noindent Where, $\mathbf{c}^{\dagger}=(\mathbf{c}^*)^T$.
+\end{align}
+ Where, $\mathbf{c}^{\dagger}=(\mathbf{c}^*)^T$.
 
@@
 $|\mathbf{k}\rangle$ has the following properties,
-\begin{eqnarray}
+\begin{align}
  \langle\mathbf{k}|S|\mathbf{k}\rangle &=&\int_{\Omega} d^3x\\
  \langle\mathbf{k}^{\prime}|S|\mathbf{k}\rangle &\simeq& 0
-\end{eqnarray}
+\end{align}
 Here, $\int_{\Omega} d^3x$ is the volume of the solution space.
@@
 \end{equation}
-\noindent Where, $\psi$ is an eigenstate of $H$ of energy $E_0$. Thus, we define the Green's function as,
+ Where, $\psi$ is an eigenstate of $H$ of energy $E_0$. Thus, we define the Green's operator as,
 \begin{equation}\label{greenfemdef}
@@
 \end{equation}
-\noindent Where, $\mathbf{I}$ is the identity matrix. The free Greens function is defined similarly by setting
+ Where, $\mathbf{I}$ is the identity matrix. The free Greens function is defined similarly by setting
 $H=B/2$. We will ultimately be testing the convergence of, $\langle\mathbf{k}|G_{fem}(z)|\mathbf{k}\rangle$.
@@
 \subsection{Modification of the Hamiltonian}
-In this section we use the techniques outlined in \cite{KukPom76}. It this paper, the Born
+In this section we use the techniques introduced in \cite{KukPom76} and elaborated in
+\cite{KukPom78}. It this paper, the Born
 series is rearranged by using orthogonally projecting pseudopotentials (OPP). It is then
@@
 \end{equation}
-\noindent The rate of convergence of the Born series is controlled by the parameter $\lambda$.
+ The rate of convergence of the Born series is controlled by the parameter $\lambda$.
 
@@
 
-\begin{eqnarray}
+\begin{align}
  \widetilde{H}&=&H+\Gamma\\
  \widetilde{H}_0&=&H_0+\Gamma
-\end{eqnarray}
+\end{align}
 
@@
 
-\noindent Physically, this means that the modified Hamiltonian has the same continuous spectrum but the discrete
+ Physically, this means that the modified Hamiltonian has the same continuous spectrum but the discrete
 spectrum has been shifted by $\lambda$.
@@
 We now use the above methods in an example with scattering off a potential in 1D. Due to the simplicity of the problem,
-it is possible to explicitly calculate the exact Green's function by inverting the appropriate matrix according the formal
-definition given in \eqref{GFdefinition}. We will examine the problem in three cases: 
+it is possible to explicitly calculate the exact Green's operator by inverting the appropriate matrix according the formal
+definition given in \eqref{GFdefinition}. We will examine the problem in two cases: 
 \begin{itemize}
  \item In the first case, we have convergence with both the modified and unmodified Hamiltonian but at a faster rate with the modification.
- \item In the second case, the unmodified case converges for certain energies and diverges for others.With the modification, we always have
-       faster convergence.
- \item In the third case, the modified hamiltonian gives a convergent Born series whereas the unmodified one diverges. 
+ \item In the second case, the modified hamiltonian gives a convergent Born series whereas the unmodified one diverges. 
 \end{itemize}
@@
 The energies of the free wavefunctions are $E_n = n^2\pi^2/200$ for $n=0,\ldots,19$. The factor
-of $200$ is a result of the discretization. The Green's function, $G(z)$ is calculated for $z=E_n+i\epsilon$,
+of $200$ is a result of the discretization. The Green's operator, $G(z)$ is calculated for $z=E_n+i\epsilon$,
 where, $\epsilon$ is a small positive number defined based on the problem. The purpose of epsilon is to avoid 
@@
 \subsection{First Case}
-In the first case, we define the potential as a finite well of strength $0.1$ with support in $4 \leq x \leq 6$,
+In the first case, we define the potential as a finite well of strength $0.01$ with support in $4.5 \leq x \leq 5.5$,
 \begin{equation}\label{potentialdef1}
 V = \begin{cases}
--0.1 &\mbox{} 4\leq x\leq 6\\
+-0.01 &\mbox{} 4.5\leq x\leq 5.5\\
 0 &\mbox{ otherwise}
@@
 \end{equation}
+
 \begin{center}
-\begin{figure}[htb]
+\begin{figure}[ht]
 \includegraphics[width=360pt, height=175pt]{series1a.png}
@@
 The Born Series is calculated for $15$ iterations. The Hamiltonian is modified using the operator defined in \eqref{potmod} with
-$\lambda=1$.  We also set $\epsilon=0.1$.
+$\lambda=1$.  We also set $\epsilon=10^{-4}$.
 
@@
 to compare the modified and unmodified cases, we plot them side by side with the modified case on the left and the unmodified one
-on the right. We follow this convention in the other two cases as well.
+on the right. We follow this convention in the other next case as well.
+We see that the series converges in both cases.
 
+\newpage
 \subsection{Second Case}
-In the second case, we define the potential as, 
-\begin{equation}\label{potentialdef2}
-V = \begin{cases}
--0.5 &\mbox{} 4\leq x\leq 6\\
-0 &\mbox{ otherwise}
-\end{cases} 
+\begin{equation}\label{potentialdef3}
+V = -100\mathtt{exp}(-(x-3.5)^2/0.5)-100\mathtt{exp}(-(x-6.5)^2/0.5)
 \end{equation}
-i.e., a finite well of strength $0.5$ with support in $4 \leq x \leq 6$,
-
-The born series is calculated for $15$ iterations. We choose $\lambda=700$ and $\epsilon=10^{-8}$.
-
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series2a.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
-
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series2b.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
-
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series2c.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
-
-We see that for the original Hamiltonian, the Born series converges in some cases but doesn't in others. However,
-with the modification, we always have convergence at a faster rate.
-\subsection{Third Case}
-\begin{equation}\label{potentialdef3}
-V = -200\mathtt{exp}(-(x-3.5)^2/0.5)-200\mathtt{exp}(-(x-6.5)^2/0.5)
-\end{equation}
-
 \begin{figure}[hb]
@@
 \includegraphics[width=360pt, height=200pt]{series3a.png}
-%\caption[\textwidth]{First Case}
+\includegraphics[width=360pt, height=200pt]{series3b.png}
+\includegraphics[width=360pt, height=200pt]{series3c.png}
+\caption[\textwidth]{Gaussian Like Potential}
 \end{figure}
 \end{center}
+% 
+% \begin{center}
+% \begin{figure}[lht]
+% \includegraphics[width=360pt, height=200pt]{series3b.png}
+% %\caption[\textwidth]{First Case}
+% \end{figure}
+% \end{center}
+% 
+% \begin{center}
+% \begin{figure}[lht]
+% \includegraphics[width=360pt, height=200pt]{series3c.png}
+% %\caption[\textwidth]{First Case}
+% \end{figure}
+% \end{center}
 
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series3b.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
+We see that for some energies the Born Series in the modified case converges to the exact value but diverges in the original case. We also note that
+for certain energies, the Born Series diverges in both cases but the divergence in the unmodified case is slower.
 
-\begin{center}
-\begin{figure}[lht]
-\includegraphics[width=360pt, height=200pt]{series3c.png}
-%\caption[\textwidth]{First Case}
-\end{figure}
-\end{center}
+\section{Conclusion}
 
-We see that the Born Series in the modified case converges to the exact value but diverges in the original case.
 \bibliographystyle{plain}

### 2012-07-26 python/axis.py

@@ -203,7 +203,38 @@ class Axis:
                 
                 for k in range(0,quad_iter):
                         (a,b) = e.val(x[k])+0j
-                        c=a * fun(-param1*x[k])
+                        c=a * fun(param1*x[k])
+                        for l in range(0,iter2):
+                                v[k][l]=c[l]*w[k]
+
+                integral=v.sum(axis=0)
+                for k in range(0,iter2):
+                        invert_vec[iter1+k]=invert_vec[iter1+k]+integral[k]
+                iter1=iter1+iter2-1
+
+	fem_vec=np.dot(invert_vec,self.overlap_inv())
+	return fem_vec
+
+    def FEM_ones(self):
+
+        nelem=self.n
+
+        #Convert momentum eigenstate into finite element representation
+        fem_vec=np.zeros(int(nelem))
+        invert_vec=np.zeros(int(nelem))
+        iter1=0
+
+
+        for e in self.e:
+                (x,w) = e.quadrature(add=10)
+                (temp_a,temp_b)=e.val(x[0])
+                iter2=len(temp_a)
+                quad_iter=len(x)
+                v=np.zeros([quad_iter,iter2])
+                
+                for k in range(0,quad_iter):
+                        (a,b) = e.val(x[k])
+                        c=a * (x[k] * x[k])
                         for l in range(0,iter2):
                                 v[k][l]=c[l]*w[k]
 

### 2012-07-26 python/bornseries.py

@@ -10,100 +10,101 @@ import scipy.linalg as la
 import pickle as pickle
 from copy import *
 from potential2 import *
+
 import time
 
 from axis import *
 
-n=100
+n=600
 order=5
-bctype='x'
 
-lb=-5.
-ub=5.
+lb=-50.
+ub=50.
 
-y=Axis(bctype,n,lb,ub,'fem',order)#+Axis('x',n,lb,ub,'fem',order)#+Axis('r',2,2.,3.,'fem')
+ycos=Axis('xopen',n,lb,ub,'fem',order)
+ysin=Axis('x',n,lb,ub,'fem',order)
 
-#y.plot()
-#plt.show()
-npart=np.floor(n/(order-1))
+Bcos=np.zeros([ycos.len(),ycos.len()])
+Bsin=np.zeros([ysin.len(),ysin.len()])
 
-if bctype=='cos':
-        nelem=npart*(order-1)+1
-elif bctype=='r' or bctype=='rho':
-        nelem=npart*(order-1)
-else:
-        nelem=npart*(order-1)-1
+Vcos=np.zeros([ycos.len(),ycos.len()])
+Vsin=np.zeros([ysin.len(),ysin.len()])
 
-#Overlap Matrix
-overlap=y.overlap()
-overlap_inv=y.overlap_inv()
+V0=1
 
-#d|d matrix
-B=np.zeros([nelem,nelem])
-#Potential Matrix
-V1=np.zeros([nelem,nelem])
-V2=np.zeros([nelem,nelem])
+for n in range(len(ycos.e)):
+	i0=ycos.e[n].i0
+	i1=i0+ycos.e[n].n
+	Bcos[i0:i1,i0:i1]=Bcos[i0:i1,i0:i1]+ycos.e[n].matrix('d|d')
+	Vcos[i0:i1,i0:i1]=Vcos[i0:i1,i0:i1]+ycos.e[n].matrix('fwell', np.array([-1.,1.,V0]))
 
-#Control Parameter
-Lambda=1
+[cos_evals,cos_evecs]=la.eig(Bcos/2.,ycos.overlap())
 
-iter1=0
+for n in range(len(ysin.e)):
+	i0=ysin.e[n].i0
+	i1=i0+ysin.e[n].n
+	Bsin[i0:i1,i0:i1]=Bsin[i0:i1,i0:i1]+ysin.e[n].matrix('d|d')
[…]

### 2012-07-26 python/bornseries_verify.py

@@ -1,0 +1,89 @@
+#!/usr/bin/env python
+import sys
+
+import numpy as np
+
+import scipy
+import scipy.special as ss
+import scipy.special.orthogonal as so
+import scipy.linalg as la
+import pickle as pickle
+from copy import *
+from potential2 import *
+
+import time
+
+from axis import *
+
+n=800
+order=5
+
+lb=0.
+ub=50.
+
+y=Axis('r',n,lb,ub,'fem',order)
+
+B=np.zeros([y.len(),y.len()])
+
+V=np.zeros([y.len(),y.len()])
+
+V0=1
+
+for n in range(len(y.e)):
+	i0=y.e[n].i0
+	i1=i0+y.e[n].n
+	B[i0:i1,i0:i1]=B[i0:i1,i0:i1]+y.e[n].matrix('d|d')
+	V[i0:i1,i0:i1]=V[i0:i1,i0:i1]+y.e[n].matrix('yukawa', np.array([0.,1.,V0]))
+
+[mom_evals,mom_evecs]=la.eig(B/2.,y.overlap())
+
+[evals,evecs]=la.eig(B/2.+V,y.overlap())
+
+#Sorting eigenvalues/eigenvectors in ascending order
+perm=np.argsort(mom_evals)
+mom_evals=np.append([0],mom_evals[perm])
+mom_evecs=mom_evecs[:,perm]
+
+temp_evecs=np.zeros([y.len(),y.len()+1])
+unit_vec=y.FEM_ones()
+temp_evecs[:,0]=unit_vec
+temp_evecs[:,1:]=mom_evecs
+mom_evecs=temp_evecs
+
+perm=np.argsort(evals)
+evals=evals[perm]
+evecs=evecs[:,perm]
+
+nbound=sum(evals<0)
+
+eps = 0j
+niter=2
+nenergy=3
+
+Tmatrix_elements=np.zeros([niter+1,nenergy])+0j
+
+V_right=np.dot(y.overlap_inv(),V)
+V_left=np.dot(V,y.overlap_inv())
+
+Vmod_right=np.dot(y.overlap_inv(),V)
+Vmod_left=np.dot(V,y.overlap_inv())
+
+Bmod = np.dot(B/2.,y.overlap_inv())
+Borig_mod = np.dot(B/2.+V,y.overlap_inv())
+
+for k in range(nenergy):		
[…]

### 2012-07-26 python/integral_nd.py

@@ -59,4 +59,4 @@ def test():
     print 'd=1',integral_nd(np.array([[0.,1.]]),func0,nquad=12)
     print 'd=2',integral_nd(np.array([[0.,1.],[0.,1.]]),func01,nquad=12)
 
-#test()
+test()

### 2012-07-26 python/periodic_finite2.py

@@ -48,6 +48,9 @@ for n in range(len(ysin.e)):
 
 [sin_evals,sin_evecs]=la.eig(Bsin/2.,ysin.overlap())
 
+[prob_cos_evals, prob_cos_evecs]=la.eig(Bcos/2.+Vcos,ycos.overlap())
+[prob_sin_evals, prob_sin_evecs]=la.eig(Bsin/2.+Vsin,ysin.overlap())
+
 #Sorting eigenvalues/eigenvectors in ascending order
 perm=np.argsort(cos_evals)
 cos_evals=cos_evals[perm]
@@ -57,33 +60,81 @@ perm=np.argsort(sin_evals)
 sin_evals=np.append([0],sin_evals[perm])
 sin_evecs=sin_evecs[:,perm]
 
+perm=np.argsort(prob_cos_evals)
+prob_cos_evals=prob_cos_evals[perm]
+prob_cos_evecs=prob_cos_evecs[:,perm]
+
+perm=np.argsort(prob_sin_evals)
+prob_sin_evals=prob_sin_evals[perm]
+prob_sin_evecs=prob_sin_evecs[:,perm]
+
+nbound_cos=sum(prob_cos_evals<0)
+nbound_sin=sum(prob_sin_evals<0)
+
 temp_evecs=np.zeros([ysin.len(),ysin.len()+1])
 temp_evecs[:,1:]=sin_evecs
 sin_evecs=temp_evecs
 
-eps = 1j
+Lambda=0
+Gam_cos=np.zeros([ycos.len(),ycos.len()])
+Gam_sin=np.zeros([ysin.len(),ysin.len()])
+
+for k in range(nbound_cos):
+	Gam_cos=Gam_cos+ycos.FEM_Outer(prob_cos_evecs[:,k],prob_cos_evecs[:,k])
+
+for k in range(nbound_sin):
+	Gam_sin=Gam_sin+ysin.FEM_Outer(prob_sin_evecs[:,k],prob_sin_evecs[:,k])
+
+eps = 0j
 niter=2
-nenergy=5
+nenergy=3
 
-Tmatrix_elements=np.zeros([niter,nenergy])
+Tmatrix_elements=np.zeros([niter+1,nenergy])+0j
 
-for k in range(5):
+V_right=np.dot(ycos.overlap_inv(),Vcos)
[…]

### 2012-07-26 python/recursive_integral.py

@@ -4,6 +4,7 @@ import numpy as np
 import scipy.special.orthogonal as so
 from copy import *
 from integral_nd import *
+from my_constants import *
 
 max_level=10      # allow recursions only to this level
 
@@ -37,7 +38,7 @@ def recursive_integral(vol,integrand,params=[],acc_rel=1.e-12,acc_abs=1.e-12,
     val_sub=[]
     for i in range(2**np.shape(vol)[0]): 
         val_sub.append(integral_nd(vol_sub(vol,i),integrand,params,nquad,[]))
-    value=np.zeros(np.shape(val_sub[0]))
+    value=np.zeros(np.shape(val_sub[0])) + 0j
     for i in range(len(val_sub)): value+= val_sub[i]
 
     # convergence criterion: relative or absolute error
@@ -59,9 +60,14 @@ def test():
     def func1(x,par=[]): return np.array([x*x*x])
     def func01(x,par=[]): return np.array([1.,x[0]*x[1],x[0]/(x[1]*x[1]+1)])
 
+    def greens(x,par=[]): 
+    	kprime=x[0]
+	x1=x[1]
+	return np.array([np.exp(1j * kprime) * np.exp(-1j * kprime * x1 ) * np.exp(1j * myPi * x1)] / ((myPi*myPi)/2 - (kprime * kprime)/2 + 1j))
+#	return np.array([x1 * x1])
+
     print 'd=1',recursive_integral(np.array([[0.,1.]]),func0,nquad=8)
     print 'd=2',recursive_integral(np.array([[0.,1.],[0.,1.]]),func01,nquad=8)
+    print 'd=1',recursive_integral(np.array([[-10.,10.],[-1.,1.]]),greens,nquad=8)
 
 test()
-
-

### 2012-08-07 Thesis - Presentation/draft1.tex
Tuesday AUG 7th 2012 14:15

@@
 \hypersetup{menubordercolor={0 0 0}}
+\usepackage{float}
 %opening
@@
 \end{equation}
- Where, $\psi$ is an eigenstate of $H$ of energy $E_0$ and $z\neq E_0$ is a complex number. Thus, we define the Green's operator as,
+ Where, $\psi$ is an eigenstate of $H$ of energy $E_0$ and $z\neq E_0$ is a complex number. Thus, we define the Green's operator as the matrix,
 \begin{equation}\label{greenfemdef}
@@
 
-Let the bound states (indexed by $a$) of the Hamiltonian $H$ be denoted by $|\psi_a\rangle$.
+Let the bound states (indexed by $a$) of the Hamiltonian $H$ be denoted by $|\varphi_a\rangle$.
 The potential is modified by the operator $\Gamma$ given by,
 \begin{equation}\label{potmod}
-  \Gamma=\lambda\displaystyle\sum_a |\psi_a\rangle\langle\psi_a|
+  \Gamma=\displaystyle\sum_a |\varphi_a\rangle\langle\varphi_a|
 \end{equation}
-The rate of convergence of the Born series is controlled by the parameter $\lambda$. It is
+The above operator is multiplied by a  positive factor $\lambda$ which controls
+the rate of convergence of the Born series. It is
 seen that for weak potentials, $\lambda$ has very little influence on the convergence while
@@
 \begin{align}
- \widetilde{H}&=H+\Gamma\\
- \widetilde{H}_0&=H_0+\Gamma
+ \widetilde{H}&=H+\lambda\Gamma\\
+ \widetilde{H}_0&=H_0+\lambda\Gamma
 \end{align}
@@
 spectrum has been shifted by $\lambda$.
+\subsubsection{Calculating Original Green's Operator}
+The original Green's operator is related to the modified Green's operator through the following
+Lippmann-Schwinger equation,
+\begin{equation}\label{lippmannschwinger2}
+ G=\widetilde{G}-\lambda\widetilde{G}\Gamma G
+\end{equation}
+We can then use the fact that,
+\begin{equation}
+ \widetilde{\Gamma}\equiv G(z)\Gamma=\Gamma G(z) = \displaystyle\sum_a \frac{|\varphi_a\rangle\langle\varphi_a|}{z-E_a} 
+\end{equation}
+where, $E_a$ is the energy of the bound state $|\varphi_a\rangle$. The above operator does not diverge
+since $Re(z)>0$ and $E_a<0$. Denoting the $n^{th}$ order
+approximation of $\widetilde{G}$ in \eqref{BornSeriesModCompact} as $\widetilde{G}^{(n)}$ we get,
+\begin{equation}\label{Gapprox}
+ G^{(n)}\approx \widetilde{G}^{(n)}+\lambda\widetilde{G}^{(n)}\widetilde{\Gamma}
+\end{equation}
 
+\subsubsection{$\lambda\rightarrow\infty$ Limit}
+Before we go on to examples where we apply the orthogonal projection method, we will briefly review the behaviour of
+the above technique in the $\lambda\rightarrow\infty$ limit as described in \cite{KukPom78}. The Lippmann-Schwinger equation \eqref{lippmannschwinger2} can also be written as,
+\begin{equation}\label{lippmannschwinger3}
+ \widetilde{G}=G+\lambda G\Gamma\widetilde{G}
+\end{equation}
+Where we have obtained an iterative expression for $\widetilde{G}$ in terms of $G$. From this we immediately obtain,
+\begin{equation}
+ \widetilde{G}=G+G\Gamma\lambda\left(\Gamma\left(I-\lambda G\right)\Gamma\right)^{-1}\Gamma G
+\end{equation}
+The admixture nonorthogonal to $\Gamma$ is,
+\begin{equation}
+ \Gamma\widetilde{G}=\Gamma\left(\Gamma\left(I-\lambda G\right)\Gamma\right)^{-1}\Gamma G \displaystyle\xrightarrow[\lambda\rightarrow\infty]{} 0,
+\end{equation}
+i.e., at $\lambda\rightarrow\infty$, the $\widetilde{G}$ is orthogonal to the subspace of the bound states
+\begin{equation}
+ \widetilde{G}=G-G\Gamma\left(\Gamma G\Gamma\right)^{-1}\Gamma G
+\end{equation}
+Where the inverse is taken only in the subspace of the bound states. The above expression thus reduces to,
+\begin{equation}
+ \widetilde{G}\rightarrow \left(I-\Gamma\right)G = G-\displaystyle\sum_i\frac{|\varphi_i\rangle\langle\varphi_i|}{E-E_i}
+\end{equation}
+Effectively, this means that the poles corresponding to the bound states have been eliminated from the original
+Green's operator. This limit was not applied in the numerical examples because of the high computational 
+overhead required. The above expression would have to be calculated separately for every energy, whereas for a finite 
+$\lambda$, only one call is required for all scattering energies.
 \section{Example Problem - Scattering in 1D}
@@
 number which is larger than the energy spacing between the bound states close to the origin.
-
 \subsection{First Case}
-%\begin{center}
-\begin{figure}
-\includegraphics[width=320pt, height=145pt]{series1a.png}
-\includegraphics[width=320pt, height=145pt]{series1b.png}
-\includegraphics[width=320pt, height=145pt]{series1c.png}
-\caption[\textwidth]{First Case - Weak Finite Well}
-\end{figure}
-%\end{center}
-
 In the first case, we define the potential as a finite well of strength $0.01$ with support in $4.5 \leq x \leq 5.5$,
@@
 
-We now plot the elements of the Born Series against the number of iterations. In order
+\subsubsection{Description of Plot}
+We now plot the elements of the Born Series against the number of iterations for different scattering energies, $E_n$. In order
 to compare the modified and unmodified cases, we plot them side by side with the modified case on the left and the unmodified one
-on the right. We follow this convention in the other next case as well.
+on the right. The x-axis denotes the number of iterations of the Born series, and the y-axis the matrix element of the
+Green's operator, $\langle\mathbf{k}|G(E_n+i\epsilon)|\mathbf{k}\rangle$ ($|\mathbf{k}\rangle$ defined in \eqref{cosevec}) in log-scale. 
+We follow this convention in the next case as well.
+
+\begin{figure}[hb]
+\includegraphics[width=320pt, height=135pt]{series1a.png}
+\includegraphics[width=320pt, height=135pt]{series1d.png}
+\end{figure}
+\begin{figure}[ht]
+%\includegraphics[width=320pt, height=135pt]{series1a.png}
+%\includegraphics[width=320pt, height=135pt]{series1d.png}
+\includegraphics[width=320pt, height=140pt]{series1b.png}
+\includegraphics[width=320pt, height=140pt]{series1c.png}
+\caption[\textwidth]{First Case - Weak Finite Well}
+\end{figure}
 
 \subsection{Second Case}
-The potential is give by,
-\begin{equation}\label{potentialdef3}
-V = -100\mathtt{exp}(-(x-3.5)^2/0.5)-100\mathtt{exp}(-(x-6.5)^2/0.5)
-\end{equation}
-We set $\epsilon=10^{-6}$ and $\lambda=100$.
-\begin{figure}[hb]
+The second test potential is a mixture of a finite-well and a gaussian. Its plot is 
+given below. We set $\epsilon=10^{-6}$ and $\lambda=100$.
+\begin{figure}[H]
 \centering
-\includegraphics[width=225pt, height=150pt]{potential1d2.png}
+\includegraphics[width=215pt, height=135pt]{potential1d2.png}
 \caption[\textwidth]{Test Potential 2}
 \end{figure}
-\begin{figure}
-\includegraphics[width=320pt, height=145pt]{series3a.png}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=320pt, height=135pt]{series3a.png}
+%\end{figure}
 \begin{figure}[ht]
-%\includegraphics[width=320pt, height=145pt]{series3a.png}
-\includegraphics[width=320pt, height=145pt]{series3b.png}
-\includegraphics[width=320pt, height=145pt]{series3c.png}
-\caption[\textwidth]{Gaussian Like Potential}
+\includegraphics[width=320pt, height=135pt]{series3a.png}
+\includegraphics[width=320pt, height=135pt]{series3b.png}
+\includegraphics[width=320pt, height=135pt]{series3c.png}
+\includegraphics[width=320pt, height=135pt]{series3d.png}
+\caption[\textwidth]{Finite Well + Gaussian}
 \end{figure}
 
+\chapter{Conclusion}
 We see that for some energies the Born Series in the modified case converges to the exact value but diverges in the original case. We also note that
 for certain energies, the Born Series diverges in both cases but the divergence in the unmodified case is slower.
-
-\chapter{Conclusion}
-
-\appendix
-\addcontentsline{toc}{chapter}{Appendices}
+%\appendix
+%\addcontentsline{toc}{chapter}{Appendices}
 %\chapter{Calculation of The Scattering Cross-section}\label{crosssection}
-\chapter{Python Implementation}\label{pythonimp}
-The main object of interest is the \verb=Axis= class, which is defined in the file \verb=axis.py=.
+%\chapter{Python Implementation}\label{pythonimp}
+%The main object of interest is the \verb=Axis= class, which is defined in the file \verb=axis.py=.
 
